### 2018-08-01 article.tex
initial files created

\RequirePackage[l2tabu,orthodox]{nag}
\documentclass[a4paper]{article}
\bibliographystyle{plain}

\pdfoutput=1

\usepackage{rwd-drafting}

\input{preamble.tex}
%%% some diagrams we won;t need in this papaer
%\input{tikzpreamble.tex}
%\input{quantumcircuits.tex}

\input{front-matter.tex}

\begin{document}
% For article.cls abstract goes here 
% other classes may have it in front-matter
\maketitle
\input{abstract.tex}

\section{Introduction}
\label{sec:intro}

\section{Mathematical Preliminaries}
\label{sec:prelim}

\section{Rewriting systems as 2-PROPS}
\label{sec:rewriting}

\section{\mathbb{Q}-valued rewriting systems}

\section{Conclusions and further work}
\label{sec:conclusion}

\small
\bibliography{all}

\end{document}

### 2018-08-02 article.tex
fixed typo
Mathematical Preliminaries: Quantale / Q-enriched category / Monoidal Q-enriched category

@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%        QUANTALE DEFINITION        %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Quantale}
+
+Let's define the following quantale $\mathbb{Q}$ over a \emph{group}, which is a monoidal category
+
+\begin{itemize}
+\item objects are elements of a lattice $L$
+\item for all objects $(x, y)$ there is a morphism between $x$ and $y$ iff $x \succeq y$ in $L$.
+\item a unit objet $e \in L$
+\item a monoidal operator $\oplus: Q \times Q \to Q$
+\end{itemize}
+
+\noindent such that for all $x, y \in L$ and $\{x_i\}_i, \{y_i\}_i \subseteq L$
+\begin{align}
+x \oplus \bigvee y_i &= \bigvee x \oplus y_i \cr
+\left(\bigvee x_i\right) \oplus y &= \bigvee x_i \oplus y
+\end{align}
+
+We require that for all objects $x$ of $\mathbb{Q}$, it exist an object $x^{-1}$ such that
+\begin{align}
+x \oplus x^{-1} &= e \cr
+x^{-1} \oplus x &= e
+\end{align}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%        Q-ENRICHED CATEGORY DEFINITION        %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{$\mathbb{Q}$-enriched category}
+
+A $\mathbb{Q}$-enriched category \catE is
+
+\begin{itemize}
+\item A set of objects
+\item For all objects $(a, b)$, $\catE(a, b) \in \mathbb{Q}$
+\end{itemize}
+
+such that for all object $a$
+$$
+e \succeq \catE(a, a)
+$$
+for all objects (a, b, c)
+$$
+\catE(b, c) \oplus \catE(a, b) \succeq \catE(a, c)
+$$
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%        MONOIDAL Q-ENRICHED CATEGORY DEFINITION        %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Monoidal $\mathbb{Q}$-enriched category}
+
+$(E, \otimes)$ is strictly monoidal if
+$$
+\catE(a, b) \oplus \catE(a', b') \succeq \catE(a \otimes a', b \otimes b')
+$$
+and symmetric if
+$$
+e \succeq \catE(a \otimes b, b \otimes a)
+$$
+
 \section{Rewriting systems as 2-PROPS}
@@
 
-\section{\mathbb{Q}-valued rewriting systems}
+\section{$\mathbb{Q}$-valued rewriting systems}
 

### 2018-08-02 article.tex
Mathematical Preliminaries: 2-category

@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%        2-CATEGORY DEFINITION        %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{2-Category}
+A 2-category \catC is a category \emph{enriched over \textbf{Cat}}, the category of categories.
+
+\begin{itemize}
+\item objects are natural number $\mathbb{N}$, called \emph{0-cells}
+\item for all 0-cells $(a, b)$, a category $\catC(a, b)$ with
+\begin{itemize}
+\item objects $f: a \to b$, called \emph{1-cells}
+\item morphisms are called \emph{2-cells}
+\item for all 2-cells $r: f \Rightarrow g$ and $s: g \Rightarrow h$, a composition operator $r \bullet s: f \Rightarrow h$ called \emph{vertical composition}
+\end{itemize}
+\item for all 0-cells $a$ and all 1-cells $f: a \to a \in \catC(a, a)$, there is an identity 2-cell $r: f \Rightarrow f$, written $Id_f$
+\item for all 0-cells $(a, b, c)$ a functor $\circ: \catC(b, c) \times \catC(a, b) \to \catC(a, c)$ called \emph{horizontal composition}
+\end{itemize}
+
+In the following we'll always write
+\begin{itemize}
+\item $a, b, c\ldots$ for \emph{0-cells}
+\item $f, g, h\ldots$ for \emph{1-cells}
+\item $r, s, t\ldots$ for \emph{2-cells}
+\end{itemize}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-Let's define the following quantale $\mathbb{Q}$ over a \emph{group}, which is a monoidal category
+A quantale $\mathbb{Q}$, over a \emph{group}, is a monoidal category
 

### 2018-08-02 article.tex
Mathematical Preliminaries: Monoidal 2-category / Interchange law

@@
 \subsection{2-Category}
-A 2-category \catC is a category \emph{enriched over \textbf{Cat}}, the category of categories.
+
+A \emph{strict} 2-category \catC is a category \emph{enriched over \textbf{Cat}}, the category of categories.
 
@@
 \item for all 0-cells $a$ and all 1-cells $f: a \to a \in \catC(a, a)$, there is an identity 2-cell $r: f \Rightarrow f$, written $Id_f$
-\item for all 0-cells $(a, b, c)$ a functor $\circ: \catC(b, c) \times \catC(a, b) \to \catC(a, c)$ called \emph{horizontal composition}
+\item for all 0-cells $(a, b, c)$ a bifunctor $\circ: \catC(b, c) \times \catC(a, b) \to \catC(a, c)$ called \emph{horizontal composition}
 \end{itemize}
 
-In the following we'll always write
+In the following we'll always write $a, b, c$ for \emph{0-cells}, $f, g, h$ for \emph{1-cells} and $r, s, t$ for \emph{2-cells}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%        MONOIDAL 2-CATEGORY DEFINITION        %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Monoidal 2-category}
+
+A \emph{strict} monoidal 2-category $(\catC, \otimes)$ is a strict 2-category with
+
 \begin{itemize}
-\item $a, b, c\ldots$ for \emph{0-cells}
-\item $f, g, h\ldots$ for \emph{1-cells}
-\item $r, s, t\ldots$ for \emph{2-cells}
+\item a 2-bifunctor $\otimes: \catC \times \catC \to \catC$ such that on 2-cells $r, s, t, u$
+$$
+(r \otimes s) \circ (t \otimes u) = (r \circ t) \otimes (s \circ u)
+$$
 \end{itemize}
+
+\begin{remark}[Interchange law]
+Because $\circ$ and $\otimes$ are bifunctor, for all 0-cells $a, b \in \catC$ and all 2-cells $r, s, t, u \in \catC(a, b)$
+
+\begin{align}
+(r \circ s) \bullet (t \circ u) &= (r \bullet t) \circ (s \bullet u) \cr
+(r \otimes s) \bullet (t \otimes u) &= (r \bullet t) \otimes (s \bullet u)
+\end{align}
+\end{remark}
 

### 2018-08-02 article.tex
Typo + Rewriting systems as 2-PROPS: Definition / Proposition 1

@@
 \item a 2-bifunctor $\otimes: \catC \times \catC \to \catC$ such that on 2-cells $r, s, t, u$
-$$
+\[
 (r \otimes s) \circ (t \otimes u) = (r \circ t) \otimes (s \circ u)
-$$
+\]
 \end{itemize}
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%        INTERCHANGE LAW REMARK        %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{remark}[Interchange law]
@@
 such that for all object $a$
-$$
+\[
 e \succeq \catE(a, a)
-$$
+\]
 for all objects (a, b, c)
-$$
+\[
 \catE(b, c) \oplus \catE(a, b) \succeq \catE(a, c)
-$$
+\]
 
@@
 $(E, \otimes)$ is strictly monoidal if
-$$
+\[
 \catE(a, b) \oplus \catE(a', b') \succeq \catE(a \otimes a', b \otimes b')
-$$
+\]
 and symmetric if
-$$
+\[
 e \succeq \catE(a \otimes b, b \otimes a)
-$$
+\]
 
@@
 \label{sec:rewriting}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%        REWRITING SYSTEM        %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+A \emph{groupoid} is a category where morphisms are invertible.
+Let's define the following strict monoidal 2-groupoid \catC
+
+\begin{itemize}
+\item 0-cells are natural number $\mathbb{N}$
+\item 1-cells are quantum circuits freely generated by $\circ$, $\otimes$ and $\{f_i\}_i$, a set of quantum circuits
+\item 2-cells are quantum circuit transformations freely generated by $\circ$, $\otimes$ and $\{r_i\}_i$, a set of ransformations, such that $\forall r: f \Rightarrow g \in \{r_i\}_i, \exists r^{-1}: g \Rightarrow f \in \{r_i\}_i$
+\[
+r \circ r^{-1} = id_f
+\]
+\item the vertical composition is the usual transformations composition
+\item the horizontal composition is the usual circuits composition
+\item the 2-bifunctor is
+\begin{itemize}
+\item on 0-cell, $a \otimes b = a + b$
+\item on $\catC(a, b)$, a functor $C(a, b) \otimes C(a', b') \to C(a + a', b + b')$
+\end{itemize}
+\end{itemize}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%        REWRITING SYSTEM PROPOSITION 1        %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{proposition}
+For all transformation $r: f \Rightarrow g$, $r$ can be writen
+\[
+r = r_1 \bullet r_2 \bullet r_3 \bullet \cdots \bullet r_i
+\]
+
+\begin{proof}
+We'll prove it by induction of the size of the transformation using the properties of the bifunctors.
+With a single transformation, we already have
+\[
+r = r_1
+\]
+For the inductive step let's assume $r$ is
+\[
+r = r_1 \bullet r_2 \bullet r_3 \bullet \cdots \bullet s
+\]
+If $s$ is such that
+\[
+s = (s_1 \bullet s_2) \circ (s_3 \bullet s_4)
+\]
+Then using the interchange law we can write
+\[
+s = (s_1 \circ s_3) \bullet (s_2 \circ s_4)
+\]
+So we have 
+\begin{align}
+r &= r_1 \bullet r_2 \bullet r_3 \bullet \cdots \bullet (s_1 \circ s_3) \bullet (s_2 \circ s_4) \cr
+r &= r_1 \bullet r_2 \bullet r_3 \bullet \cdots \bullet r_{i - 1} \bullet r_i
+\end{align}
+Otherwise we can always add an identity transformation to get the interchange law.
+\begin{align}
+s_1 \circ (s_3 \bullet s_4) &\Rightarrow (s_1 \bullet id) \circ (s_2 \bullet s_3) \cr
+(s_1 \bullet s_2) \circ s_3 &\Rightarrow (s_1 \bullet s_2) \circ (id \bullet s_3)
+\end{align}
+The same is true with
+\[
+s = (s_1 \bullet s_2) \otimes (s_3 \bullet s_4)
+\]
+\end{proof}
+\end{proposition}
 

### 2018-08-02 article.tex
Rewriting systems as 2-PROPS: Proposition 2 / Proposition 2 corollary
Q-valued rewriting systems: Homomorphism / Remark
Q-valued rewriting systems: Proposition / Proposition corollary

@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%        REWRITING SYSTEM PROPOSITION 2        %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{proposition}
+For all transformation $r: f \Rightarrow g$, $r$ can be writen
+\begin{align}
+r &= r_1 \circ r_2 \circ r_3 \circ \cdots \circ r_j \cr
+r &= r_1 \otimes r_2 \otimes r_3 \otimes \cdots \otimes r_k
+\end{align}
+
+\begin{proof}
+The proof is the same but this time we have to use
+\[
+(r \otimes s) \circ (t \otimes u) = (r \circ t) \otimes (s \circ u)
+\]
+\end{proof}
+\end{proposition}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%        REWRITING SYSTEM PROPOSITION 2 COROLLARY        %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{corollary}
+For all transformation $r: f \Rightarrow g$, $r$ can be writen
+\[
+r = r_1 \bullet r_2 \bullet r_3 \bullet \cdots \bullet r_i
+\]
+Such that for all $i$
+\[
+r_i = r_{i, 1} \circ r_{i, 2} \circ r_{i, 3} \circ \cdots \circ r_{i, j}
+\]
+and for all $j$
+\[
+r_{i, j} = \bigotimes_k^l r_{i, j, k}
+\]
+a tensor products of transformations.
+\end{corollary}
+
 \section{$\mathbb{Q}$-valued rewriting systems}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%        HOMOMORPHISM        %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{$\catC(a, b)$ to $\mathbb{Q}$ homomorphism}
+
+Let's define the following homomorphism $F$ from $\catC(a, b)$ to $\mathbb{Q}$. For all $f: a \to b \in \{f_i\}_i$
+\[
+F(f) \in L
+\]
+For all $r: f \Rightarrow g \in \{r_i\}_i$
+\[
+F(r) = F(f) \oplus F(g)^{-1}
+\]
+
+On 1-cells
+\begin{align}
+F(f \circ g) &\preceq F(f) \oplus F(g) \cr
+F(f \otimes g) &\preceq F(f) \oplus F(g)
+\end{align}
+
+On 2-cells
+\begin{align}
+F(r \circ s) &= F(r) \oplus F(s) \cr
+F(r \otimes s) &= F(r) \oplus F(s) \cr
+F(r \bullet s) &= F(r) \oplus F(s)
+\end{align}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%        HOMOMORPHISM REMARK        %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{remark}
+Note that because
+\[
+F(r: f \Rightarrow g) = F(f) \oplus F(g)^{-1}
+\]
+then
+\begin{align}
+F(r: f \Rightarrow f) &= F(f) \oplus F(f)^{-1} \cr
+&= e
+\end{align}
+that is
+\begin{align}
+\forall f, F(id_f) = e
+\end{align}
+\end{remark}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%        HOMOMORPHISM PROPOSITION        %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{proposition}
+Given two transformations $r, s: f \Rightarrow g$, then $F(r) = F(s)$.
+
+\begin{proof}
+The set of all possible transformation is freely generated by $\{r_i\}_i$, $\bullet$, $\circ$ and $\otimes$ so both $r$ and $s$ can be written
+\begin{align}
+r &= r_1 \odot r_2 \odot r_3 \odot \cdots \odot r_n \cr
+s &= s_1 \odot s_2 \odot s_3 \odot \cdots \odot s_m
+\end{align}
+with $\odot \in \{\bullet, \circ, \otimes\}$ and for all i
+\begin{align}
+r_i &\in \{f_i\}_i \cr
+s_i &\in \{f_i\}_i
+\end{align}
+If $n = 1$ then $r \in \{f_i\}_i$, so $F(r) = F(f) \oplus F(g)^{-1}$.
+Otherwise
+\begin{align}
+F(r) &= F(r_1 \odot r_2 \odot \cdots \odot r_n) \cr
+&= F(r_1) \oplus F(r_2) \oplus \cdots \oplus F(r_n) \cr
+&= F(f \Rightarrow h_1) \oplus F(h_1 \Rightarrow h_2) \oplus \cdots \oplus F(h_n \Rightarrow g) \cr
+&= F(f) \oplus F(h_1)^{-1} \oplus F(h_1) \oplus F(h_2)^{-1} \oplus \cdots \oplus F(h_n) \oplus F(g)^{-1}
+\end{align}
+A easy inductive proof can show us that
+\[
+F(f) \oplus F(h_1)^{-1} \oplus F(h_1) \oplus F(h_2)^{-1} \oplus F(h_2) \oplus \cdots \oplus F(h_n) \oplus F(g)^{-1} = F(f) \oplus F(g)^{-1}
+\]
+The same is true for $s$, so
+\[
+F(r) = F(s) = F(f) \oplus F(g)^{-1}
+\]
+\end{proof}
+\end{proposition}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%        HOMOMORPHISM PROPOSITION CO           %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{corollary}
+For all transformations $r: f \Rightarrow g$, then $F(r) = F(f) \oplus F(g)^{-1}$.
+Note that the definition of $F$ gives that only for $r \in \{f_i\}_i$.
+\end{corollary}
 

### 2018-08-03 article.tex
Q-valued rewriting systems: Proposition 2

@@
 \item on 0-cell, $a \otimes b = a + b$
-\item on $\catC(a, b)$, a functor $C(a, b) \otimes C(a', b') \to C(a + a', b + b')$
+\item on $\catC(a, b)$, a functor $\catC(a, b) \otimes C(a', b') \to C(a + a', b + b')$
 \end{itemize}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%        HOMOMORPHISM PROPOSITION        %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%        HOMOMORPHISM PROPOSITION 1        %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{proposition}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%        HOMOMORPHISM PROPOSITION CO           %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%        HOMOMORPHISM PROPOSITION 1 COROLLARY        %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{corollary}
@@
 \end{corollary}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%        HOMOMORPHISM PROPOSITION 2        %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{proposition}
+The following category $\catC(\mathbb{Q})$ is a $\mathbb{Q}$-category
+\begin{itemize}
+\item A set of objects which are the 1-cells of $\catC$
+\item For all objects $(f, g)$
+\end{itemize}
+\[
+\catE(f, g) = \begin{cases} {\rm if~} f \Rightarrow g \in \catC, &F(f \Rightarrow g) \cr {\rm otherwise~} &\top \end{cases}
+\]
+With $\top \in Q$ such that for all $x \in \mathbb{Q}$
+\[
+x \preceq \top
+\]
+
+\begin{proof}
+For all 1-cells $(f, g)$, if it exists a unique transformation $f \Rightarrow g$, then
+\begin{align}
+\catE(f, g) &= F(f \Rightarrow g) \in \mathbb{Q} \cr
+&= F(f) \oplus F(g)^{-1}
+\end{align}
+otherwise we already have $\catE(f, g) = \top \in \mathbb{Q}$.
+For all 1-cell $f$, we know that there is a identity transformation $f \Rightarrow f$ and
+\begin{align}
+\catE(f, f) &= F(id_f) \cr
+&= F(f \Rightarrow f)\cr
+&= F(f) \oplus F(f)^{-1} \cr
+&= e
+\end{align}
+Finally, for all 1-cells $(f, g, h)$ such that there exists $r: f \Rightarrow g$ and $s: g \Rightarrow h$ in $\catC$, then the homomorphism $F$ give us
+\begin{align}
+\catE(f, h) &= F(f \Rightarrow h) \cr
+&= F(s\circ r) \cr
+&= F(s) \oplus F(r) \cr
+&= F(g \Rightarrow h) \oplus F(f \Rightarrow g) \cr
+&= \catE(g, h) \oplus \catE(f, g)
+\end{align}
+If either $r$ or $s$ doesn't exists in $\catC$ then
+\[
+\catE(b, c) \oplus \catE(a, b) \succeq \catE(a, c)
+\]
+become
+\[\begin{matrix}
+&\catE(b, c) \oplus \top &\succeq \top \cr
+{\rm or~} &\top \oplus \catE(a, b) &\succeq \top \cr
+{\rm or~} &\top \oplus \top &\succeq \top
+\end{matrix}\]
+which is still true.
+\end{proof}
+\end{proposition}
 

### 2018-08-03 article.tex
Q-valued rewriting systems: Proposition 3 / Proposition 3 corollary

@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%        HOMOMORPHISM PROPOSITION 3        %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{proposition}
+$\catC(\mathbb{Q})$ is strictly symmetric monoidal with $\otimes$, the 2-functor of $\catC$.
+
+\begin{proof}
+The monoidal structure is immediate using the fact that
+\[
+F(r) \oplus F(s) \succeq F(r \otimes s)
+\]
+that is
+\[
+F(r: f \Rightarrow g) \oplus F(s: f' \Rightarrow g') = F(r \otimes s: f \otimes f' \Rightarrow g \otimes g')
+\]
+so
+\[
+\catE(f, g) \oplus \catE(f', g') = \catE(f \otimes f', g \otimes g')
+\]
+For the symmetry, we use the fact that $F(f \Rightarrow g) \oplus F(g \Rightarrow f) = e$
+\begin{align}
+\catE(f \otimes g, g \otimes f) &= F(f \otimes g \Rightarrow g \otimes g) \cr
+&= F(f \Rightarrow g \otimes g \Rightarrow f) \cr
+&\preceq F(f \Rightarrow g) \oplus F(g \Rightarrow f) \cr
+&= F(f) \oplus F(g)^{-1} \oplus F(g) \oplus F(f)^{-1} \cr
+&= e
+\end{align}
+\end{proof}
+\end{proposition}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%        HOMOMORPHISM PROPOSITION 3 COROLLARY        %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{corollary}
+Given $s: f \Rightarrow g$ a transformation, and $s^{-1}: g \Rightarrow f$ the inverse transformation
+\[
+F(s \otimes s^{-1}) = e
+\]
+\end{corollary}
+
 \section{Conclusions and further work}

### 2018-08-06 article.tex
edits to beginning of preliminaries; moved quantale stuff
later

@@
 
+We assume that the reader is familiar with the notion of symmetric
+monoidal category; a standard reference is
+\cite{MacLane:CatsWM:1971}.
+
+\begin{definition}\label{def:PROP}
+  A \emph{PROP} is strict symmetric monoidal category whose objects
+  are the natural numbers, and where the action of the tensor product on
+  objects is given by addition.
+\end{definition}
+
+PROPs have been used to formalise many algebraic structures
+\cite{Lack:2004sf}, however our interest here is to formalise a
+general notion of \emph{circuit} as a free PROP over a certain set of
+gates.
+
+\begin{example}  \label{ex:boolean-circs}
+  TODO -- boolean circuits
+\end{example}
+
+\begin{example}  \label{ex:quantum-circs}
+  TODO -- quanutm circuits
+\end{example}
+
+\begin{example}  \label{ex:zx-calculus}
+  TODO -- zx-calculus
+\end{example}
+
+\noindent
+Note that all of these examples usually include an equational theory.
+We will don't include these equations in the  definition since we will
+model them as higher dimensional arrows.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{2-Category}
 
-A \emph{strict} 2-category \catC is a category \emph{enriched over \textbf{Cat}}, the category of categories.
+\begin{definition}  \label{def:strict-two-cat}
+A \emph{strict} 2-category \catC is a category \emph{enriched over
+  \textbf{Cat}}, the category of categories, \emph{viz:}
+\begin{itemize}
+\item A category \catC, whose objects and arrows are known as
+  0-cells and 1-cells respectively.
+\item For every pair of objects $A$ and $B$, we have a
+  category $\catC(A, B)$, whose arrows are known as 2-cells.
+\item For 2-cells $\alpha : f \Rightarrow g$ and $\beta : g
+  \Rightarrow h$ in $\catC(A,B)$ their (vertical) composite is denoted
+  $\beta\bullet\alpha : f \Rightarrow h$.
+\item For every 1-cell $f:A\to B$ we denote its identity 2-cell by
+  $Id_f:f \Rightarrow f$.
+\item For all 0-cells $a$, $b$, $c$ a bifunctor $\circ: \catC(b, c)
+  \times \catC(a, b) \to \catC(a, c)$ called \emph{horizontal
+    composition} of 2-cells.
+\end{itemize}
+\REM{Need here the name for a 2-cat with all 2-cells invertible;  the
+  term 2-groupoid refers to the case where both 1- and 2-cells are
+  invertible.}
+\end{definition}
 
-\begin{itemize}
-\item objects are natural number $\mathbb{N}$, called \emph{0-cells}
-\item for all 0-cells $(a, b)$, a category $\catC(a, b)$ with
-\begin{itemize}
-\item objects $f: a \to b$, called \emph{1-cells}
-\item morphisms are called \emph{2-cells}
-\item for all 2-cells $r: f \Rightarrow g$ and $s: g \Rightarrow h$, a composition operator $r \bullet s: f \Rightarrow h$ called \emph{vertical composition}
-\end{itemize}
-\item for all 0-cells $a$ and all 1-cells $f: a \to a \in \catC(a, a)$, there is an identity 2-cell $r: f \Rightarrow f$, written $Id_f$
-\item for all 0-cells $(a, b, c)$ a bifunctor $\circ: \catC(b, c) \times \catC(a, b) \to \catC(a, c)$ called \emph{horizontal composition}
-\end{itemize}
+\noindent
+In the following we'll always write $A, B, C$ for \emph{0-cells}, $f,
+g, h$ for \emph{1-cells} and $\alpha$, $\beta$, $\gamma$ for
+\emph{2-cells}.
 
-In the following we'll always write $a, b, c$ for \emph{0-cells}, $f, g, h$ for \emph{1-cells} and $r, s, t$ for \emph{2-cells}.
+\begin{definition}
+  \label{def:two-functor}
+  DEFINITION OF 2-FUNCTOR here
+\end{definition}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Monoidal 2-category}
 
-A \emph{strict} monoidal 2-category $(\catC, \otimes)$ is a strict 2-category with
-
+\begin{definition}\label{def:monoidal-two-cat}
+A \emph{strict} monoidal 2-category $(\catC, \otimes)$ is a strict
+2-category with
+\REM{Is this \emph{required}?  Doesn't it follow from the  definition
+  of bifunctor?  (It does in the 1-category case, so it should do
+  here!)}
 \begin{itemize}
-\item a 2-bifunctor $\otimes: \catC \times \catC \to \catC$ such that on 2-cells $r, s, t, u$
+\item a 2-bifunctor $\otimes: \catC \times \catC \to \catC$ such that
+  on 2-cells $r, s, t, u$
 \[
@@
 \end{itemize}  
+\end{definition}
 
@@
 \begin{remark}[Interchange law]
-Because $\circ$ and $\otimes$ are bifunctor, for all 0-cells $a, b \in \catC$ and all 2-cells $r, s, t, u \in \catC(a, b)$
-
+Since $\circ$ and $\otimes$ are bifunctorial, we have the following
+interchange equations for all 0-cells $A, B \in \catC$ and all 2-cells $r, s, t, u \in \catC(A, B)$
 \begin{align}
-(r \circ s) \bullet (t \circ u) &= (r \bullet t) \circ (s \bullet u) \cr
+(r \circ s) \bullet (t \circ u) &= (r \bullet t) \circ (s \bullet u) \\
 (r \otimes s) \bullet (t \otimes u) &= (r \bullet t) \otimes (s \bullet u)
@@
 \end{remark}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%        QUANTALE DEFINITION        %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Quantale}
-
-A quantale $\mathbb{Q}$, over a \emph{group}, is a monoidal category
-
-\begin{itemize}
-\item objects are elements of a lattice $L$
-\item for all objects $(x, y)$ there is a morphism between $x$ and $y$ iff $x \succeq y$ in $L$.
-\item a unit objet $e \in L$
-\item a monoidal operator $\oplus: Q \times Q \to Q$
-\end{itemize}
-
-\noindent such that for all $x, y \in L$ and $\{x_i\}_i, \{y_i\}_i \subseteq L$
-\begin{align}
-x \oplus \bigvee y_i &= \bigvee x \oplus y_i \cr
-\left(\bigvee x_i\right) \oplus y &= \bigvee x_i \oplus y
-\end{align}
-
-We require that for all objects $x$ of $\mathbb{Q}$, it exist an object $x^{-1}$ such that
-\begin{align}
-x \oplus x^{-1} &= e \cr
-x^{-1} \oplus x &= e
-\end{align}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%        Q-ENRICHED CATEGORY DEFINITION        %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{$\mathbb{Q}$-enriched category}
-
-A $\mathbb{Q}$-enriched category \catE is
-
-\begin{itemize}
-\item A set of objects
-\item For all objects $(a, b)$, $\catE(a, b) \in \mathbb{Q}$
-\end{itemize}
-
-such that for all object $a$
-\[
-e \succeq \catE(a, a)
-\]
-for all objects (a, b, c)
-\[
-\catE(b, c) \oplus \catE(a, b) \succeq \catE(a, c)
-\]
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%        MONOIDAL Q-ENRICHED CATEGORY DEFINITION        %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Monoidal $\mathbb{Q}$-enriched category}
-
-$(E, \otimes)$ is strictly monoidal if
-\[
-\catE(a, b) \oplus \catE(a', b') \succeq \catE(a \otimes a', b \otimes b')
-\]
-and symmetric if
-\[
-e \succeq \catE(a \otimes b, b \otimes a)
-\]
 
@@
 \end{corollary}
+
+\section{Quantale Enriched Categories}
+\label{sec:quant-enrich-categ}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%        QUANTALE DEFINITION        %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Quantale}
+
+A quantale $\mathbb{Q}$, over a \emph{group}, is a monoidal category
+
+\begin{itemize}
+\item objects are elements of a lattice $L$
+\item for all objects $(x, y)$ there is a morphism between $x$ and $y$ iff $x \succeq y$ in $L$.
+\item a unit objet $e \in L$
+\item a monoidal operator $\oplus: Q \times Q \to Q$
+\end{itemize}
+
+\noindent such that for all $x, y \in L$ and $\{x_i\}_i, \{y_i\}_i \subseteq L$
+\begin{align}
+x \oplus \bigvee y_i &= \bigvee x \oplus y_i \cr
+\left(\bigvee x_i\right) \oplus y &= \bigvee x_i \oplus y
+\end{align}
+
+We require that for all objects $x$ of $\mathbb{Q}$, it exist an object $x^{-1}$ such that
+\begin{align}
+x \oplus x^{-1} &= e \cr
+x^{-1} \oplus x &= e
+\end{align}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%        Q-ENRICHED CATEGORY DEFINITION        %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{$\mathbb{Q}$-enriched category}
+
+A $\mathbb{Q}$-enriched category \catE is
+
+\begin{itemize}
+\item A set of objects
+\item For all objects $(a, b)$, $\catE(a, b) \in \mathbb{Q}$
+\end{itemize}
+
+such that for all object $a$
+\[
+e \succeq \catE(a, a)
+\]
+for all objects (a, b, c)
+\[
+\catE(b, c) \oplus \catE(a, b) \succeq \catE(a, c)
+\]
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%        MONOIDAL Q-ENRICHED CATEGORY DEFINITION        %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Monoidal $\mathbb{Q}$-enriched category}
+
+$(E, \otimes)$ is strictly monoidal if
+\[
+\catE(a, b) \oplus \catE(a', b') \succeq \catE(a \otimes a', b \otimes b')
+\]
+and symmetric if
+\[
+e \succeq \catE(a \otimes b, b \otimes a)
+\]
 

### 2018-08-06 article.tex
typo and small modifications

--- 2018-08-03 article.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{2-Category}
-
+\begin{definition}[2-Category]
 A \emph{strict} 2-category \catC is a category \emph{enriched over \textbf{Cat}}, the category of categories.
@@
 In the following we'll always write $a, b, c$ for \emph{0-cells}, $f, g, h$ for \emph{1-cells} and $r, s, t$ for \emph{2-cells}.
+\end{definition}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Monoidal 2-category}
-
+\begin{definition}[Monoidal 2-category]
 A \emph{strict} monoidal 2-category $(\catC, \otimes)$ is a strict 2-category with
@@
 \end{itemize}
+\end{definition}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Quantale}
-
-A quantale $\mathbb{Q}$, over a \emph{group}, is a monoidal category
-
-\begin{itemize}
-\item objects are elements of a lattice $L$
-\item for all objects $(x, y)$ there is a morphism between $x$ and $y$ iff $x \succeq y$ in $L$.
-\item a unit objet $e \in L$
-\item a monoidal operator $\oplus: Q \times Q \to Q$
-\end{itemize}
-
-\noindent such that for all $x, y \in L$ and $\{x_i\}_i, \{y_i\}_i \subseteq L$
+\begin{definition}[Quantale $\mathbb{Q}$]
+A quantale is a \emph{lattice} with a binary operator such that for all elements of the lattice $x, y$ and $\{x_i\}_i, \{y_i\}_i$
 \begin{align}
@@
 
-We require that for all objects $x$ of $\mathbb{Q}$, it exist an object $x^{-1}$ such that
+We require that for all $x$, it exist $x^{-1}$ such that
 \begin{align}
@@
 
+Such a quantale can be seen as a monoidal category over a \emph{group}.
+\end{definition}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{$\mathbb{Q}$-enriched category}
-
+\begin{definition}[$\mathbb{Q}$-enriched category]
 A $\mathbb{Q}$-enriched category \catE is
@@
 \]
+\end{definition}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Monoidal $\mathbb{Q}$-enriched category}
-
+\begin{definition}[Monoidal $\mathbb{Q}$-enriched category]
 $(E, \otimes)$ is strictly monoidal if
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%        REWRITING SYSTEM        %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%        REWRITING SYSTEM DEFINITION        %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{definition}
 A \emph{groupoid} is a category where morphisms are invertible.
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%        HOMOMORPHISM        %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{$\catC(a, b)$ to $\mathbb{Q}$ homomorphism}
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%        HOMOMORPHISM DEFINITION        %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{definition}[$\catC(a, b)$ to $\mathbb{Q}$ homomorphism]
 Let's define the following homomorphism $F$ from $\catC(a, b)$ to $\mathbb{Q}$. For all $f: a \to b \in \{f_i\}_i$
 \[
-F(f) \in L
+V(f) \in \mathbb{Q}
 \]
@@
 \[
-F(r) = F(f) \oplus F(g)^{-1}
+V(r) = V(f) \oplus V(g)^{-1}
 \]
@@
 \begin{align}
-F(f \circ g) &\preceq F(f) \oplus F(g) \cr
-F(f \otimes g) &\preceq F(f) \oplus F(g)
+V(f \circ g) &\preceq V(f) \oplus V(g) \cr
+V(f \otimes g) &\preceq V(f) \oplus V(g)
 \end{align}
@@
 \begin{align}
-F(r \circ s) &= F(r) \oplus F(s) \cr
-F(r \otimes s) &= F(r) \oplus F(s) \cr
-F(r \bullet s) &= F(r) \oplus F(s)
+V(r \circ s) &= V(r) \oplus V(s) \cr
+V(r \otimes s) &= V(r) \oplus V(s) \cr
+V(r \bullet s) &= V(r) \oplus V(s)
 \end{align}
+\end{definition}
 
@@
 \[
-F(r: f \Rightarrow g) = F(f) \oplus F(g)^{-1}
+V(r: f \Rightarrow g) = V(f) \oplus V(g)^{-1}
 \]
@@
 \begin{align}
-F(r: f \Rightarrow f) &= F(f) \oplus F(f)^{-1} \cr
+V(r: f \Rightarrow f) &= V(f) \oplus V(f)^{-1} \cr
 &= e
@@
 \begin{align}
-\forall f, F(id_f) = e
+\forall f, V(id_f) = e
 \end{align}
@@
 \begin{proposition}
-Given two transformations $r, s: f \Rightarrow g$, then $F(r) = F(s)$.
+Given two transformations $r, s: f \Rightarrow g$, then $V(r) = V(s)$.
 
@@
 \end{align}
-If $n = 1$ then $r \in \{f_i\}_i$, so $F(r) = F(f) \oplus F(g)^{-1}$.
+If $n = 1$ then $r \in \{f_i\}_i$, so $V(r) = V(f) \oplus V(g)^{-1}$.
 Otherwise
 \begin{align}
-F(r) &= F(r_1 \odot r_2 \odot \cdots \odot r_n) \cr
-&= F(r_1) \oplus F(r_2) \oplus \cdots \oplus F(r_n) \cr
-&= F(f \Rightarrow h_1) \oplus F(h_1 \Rightarrow h_2) \oplus \cdots \oplus F(h_n \Rightarrow g) \cr
-&= F(f) \oplus F(h_1)^{-1} \oplus F(h_1) \oplus F(h_2)^{-1} \oplus \cdots \oplus F(h_n) \oplus F(g)^{-1}
+V(r) &= V(r_1 \odot r_2 \odot \cdots \odot r_n) \cr
+&= V(r_1) \oplus V(r_2) \oplus \cdots \oplus V(r_n) \cr
+&= V(f \Rightarrow h_1) \oplus V(h_1 \Rightarrow h_2) \oplus \cdots \oplus V(h_n \Rightarrow g) \cr
+&= V(f) \oplus V(h_1)^{-1} \oplus V(h_1) \oplus V(h_2)^{-1} \oplus \cdots \oplus V(h_n) \oplus V(g)^{-1}
 \end{align}
@@
 \[
-F(f) \oplus F(h_1)^{-1} \oplus F(h_1) \oplus F(h_2)^{-1} \oplus F(h_2) \oplus \cdots \oplus F(h_n) \oplus F(g)^{-1} = F(f) \oplus F(g)^{-1}
+V(f) \oplus V(h_1)^{-1} \oplus V(h_1) \oplus V(h_2)^{-1} \oplus V(h_2) \oplus \cdots \oplus V(h_n) \oplus V(g)^{-1} = V(f) \oplus V(g)^{-1}
 \]
@@
 \[
-F(r) = F(s) = F(f) \oplus F(g)^{-1}
+V(r) = V(s) = V(f) \oplus V(g)^{-1}
 \]
@@
 \begin{corollary}
-For all transformations $r: f \Rightarrow g$, then $F(r) = F(f) \oplus F(g)^{-1}$.
+For all transformations $r: f \Rightarrow g$, then $V(r) = V(f) \oplus V(g)^{-1}$.
 Note that the definition of $F$ gives that only for $r \in \{f_i\}_i$.
@@
 \[
-\catE(f, g) = \begin{cases} {\rm if~} f \Rightarrow g \in \catC, &F(f \Rightarrow g) \cr {\rm otherwise~} &\top \end{cases}
+\catE(f, g) = \begin{cases} {\rm if~} f \Rightarrow g \in \catC, &V(f \Rightarrow g) \cr {\rm otherwise~} &\top \end{cases}
 \]
@@
 \begin{align}
-\catE(f, g) &= F(f \Rightarrow g) \in \mathbb{Q} \cr
-&= F(f) \oplus F(g)^{-1}
+\catE(f, g) &= V(f \Rightarrow g) \in \mathbb{Q} \cr
+&= V(f) \oplus V(g)^{-1}
 \end{align}
@@
 \begin{align}
-\catE(f, f) &= F(id_f) \cr
-&= F(f \Rightarrow f)\cr
-&= F(f) \oplus F(f)^{-1} \cr
+\catE(f, f) &= V(id_f) \cr
+&= V(f \Rightarrow f)\cr
+&= V(f) \oplus V(f)^{-1} \cr
 &= e
@@
 \begin{align}
-\catE(f, h) &= F(f \Rightarrow h) \cr
-&= F(s\circ r) \cr
-&= F(s) \oplus F(r) \cr
-&= F(g \Rightarrow h) \oplus F(f \Rightarrow g) \cr
+\catE(f, h) &= V(f \Rightarrow h) \cr
+&= V(s\circ r) \cr
+&= V(s) \oplus V(r) \cr
+&= V(g \Rightarrow h) \oplus V(f \Rightarrow g) \cr
 &= \catE(g, h) \oplus \catE(f, g)
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%        HOMOMORPHISM PROPOSITION 3        %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{proposition}
+$\catC(\mathbb{Q})$ is strictly symmetric monoidal with $\otimes$, the 2-functor of $\catC$.
+
+\begin{proof}
+The monoidal structure is immediate using the fact that
+\[
+V(r) \oplus V(s) \succeq V(r \otimes s)
+\]
+that is
+\[
+V(r: f \Rightarrow g) \oplus V(s: f' \Rightarrow g') = V(r \otimes s: f \otimes f' \Rightarrow g \otimes g')
+\]
+so
+\[
+\catE(f, g) \oplus \catE(f', g') = \catE(f \otimes f', g \otimes g')
+\]
+For the symmetry, we use the fact that $V(f \Rightarrow g) \oplus V(g \Rightarrow f) = e$
+\begin{align}
+\catE(f \otimes g, g \otimes f) &= V(f \otimes g \Rightarrow g \otimes g) \cr
+&= V(f \Rightarrow g \otimes g \Rightarrow f) \cr
+&\preceq V(f \Rightarrow g) \oplus V(g \Rightarrow f) \cr
+&= V(f) \oplus V(g)^{-1} \oplus V(g) \oplus V(f)^{-1} \cr
+&= e
+\end{align}
+\end{proof}
+\end{proposition}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%        HOMOMORPHISM PROPOSITION 3 COROLLARY        %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{corollary}
+Given $s: f \Rightarrow g$ a transformation, and $s^{-1}: g \Rightarrow f$ the inverse transformation
+\[
+V(s \otimes s^{-1}) = e
+\]
+\end{corollary}
+
 \section{Conclusions and further work}

### 2018-08-06 article.tex

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{definition}[2-Category]
-A \emph{strict} 2-category \catC is a category \emph{enriched over \textbf{Cat}}, the category of categories.
 
+\begin{definition}  \label{def:strict-two-cat}
+A \emph{strict} 2-category \catC is a category \emph{enriched over
+  \textbf{Cat}}, the category of categories, \emph{viz:}
 \begin{itemize}
-\item objects are natural number $\mathbb{N}$, called \emph{0-cells}
-\item for all 0-cells $(a, b)$, a category $\catC(a, b)$ with
-\begin{itemize}
-\item objects $f: a \to b$, called \emph{1-cells}
-\item morphisms are called \emph{2-cells}
-\item for all 2-cells $r: f \Rightarrow g$ and $s: g \Rightarrow h$, a composition operator $r \bullet s: f \Rightarrow h$ called \emph{vertical composition}
+\item A category \catC, whose objects and arrows are known as
+  0-cells and 1-cells respectively.
+\item For every pair of objects $A$ and $B$, we have a
+  category $\catC(A, B)$, whose arrows are known as 2-cells.
+\item For 2-cells $\alpha : f \Rightarrow g$ and $\beta : g
+  \Rightarrow h$ in $\catC(A,B)$ their (vertical) composite is denoted
+  $\beta\bullet\alpha : f \Rightarrow h$.
+\item For every 1-cell $f:A\to B$ we denote its identity 2-cell by
+  $Id_f:f \Rightarrow f$.
+\item For all 0-cells $a$, $b$, $c$ a bifunctor $\circ: \catC(b, c)
+  \times \catC(a, b) \to \catC(a, c)$ called \emph{horizontal
+    composition} of 2-cells.
 \end{itemize}
-\item for all 0-cells $a$ and all 1-cells $f: a \to a \in \catC(a, a)$, there is an identity 2-cell $r: f \Rightarrow f$, written $Id_f$
-\item for all 0-cells $(a, b, c)$ a bifunctor $\circ: \catC(b, c) \times \catC(a, b) \to \catC(a, c)$ called \emph{horizontal composition}
-\end{itemize}
+\REM{Need here the name for a 2-cat with all 2-cells invertible;  the
+  term 2-groupoid refers to the case where both 1- and 2-cells are
+  invertible.}
+\end{definition}
 
-In the following we'll always write $a, b, c$ for \emph{0-cells}, $f, g, h$ for \emph{1-cells} and $r, s, t$ for \emph{2-cells}.
+\noindent
+In the following we'll always write $A, B, C$ for \emph{0-cells}, $f,
+g, h$ for \emph{1-cells} and $\alpha$, $\beta$, $\gamma$ for
+\emph{2-cells}.
+
+\begin{definition}
+  \label{def:two-functor}
+  DEFINITION OF 2-FUNCTOR here
 \end{definition}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{definition}[Monoidal 2-category]
-A \emph{strict} monoidal 2-category $(\catC, \otimes)$ is a strict 2-category with
 
+\begin{definition}\label{def:monoidal-two-cat}
+A \emph{strict} monoidal 2-category $(\catC, \otimes)$ is a strict
+2-category with
+\REM{Is this \emph{required}?  Doesn't it follow from the  definition
+  of bifunctor?  (It does in the 1-category case, so it should do
+  here!)}
 \begin{itemize}
-\item a 2-bifunctor $\otimes: \catC \times \catC \to \catC$ such that on 2-cells $r, s, t, u$
+\item a 2-bifunctor $\otimes: \catC \times \catC \to \catC$ such that
+  on 2-cells $r, s, t, u$
 \[
@@
 \]
+\REM{This definition doesn't quite work without the additoin type
+  information on the 2-cells}
 \end{itemize}
@@
 \end{remark}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%        QUANTALE DEFINITION        %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{definition}[Quantale $\mathbb{Q}$]
-A quantale is a \emph{lattice} with a binary operator such that for all elements of the lattice $x, y$ and $\{x_i\}_i, \{y_i\}_i$
-\begin{align}
-x \oplus \bigvee y_i &= \bigvee x \oplus y_i \cr
-\left(\bigvee x_i\right) \oplus y &= \bigvee x_i \oplus y
-\end{align}
-
-We require that for all $x$, it exist $x^{-1}$ such that
-\begin{align}
-x \oplus x^{-1} &= e \cr
-x^{-1} \oplus x &= e
-\end{align}
-
-Such a quantale can be seen as a monoidal category over a \emph{group}.
-\end{definition}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%        Q-ENRICHED CATEGORY DEFINITION        %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{definition}[$\mathbb{Q}$-enriched category]
-A $\mathbb{Q}$-enriched category \catE is
-
-\begin{itemize}
-\item A set of objects
-\item For all objects $(a, b)$, $\catE(a, b) \in \mathbb{Q}$
-\end{itemize}
-
-such that for all object $a$
-\[
-e \succeq \catE(a, a)
-\]
-for all objects (a, b, c)
-\[
-\catE(b, c) \oplus \catE(a, b) \succeq \catE(a, c)
-\]
-\end{definition}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%        MONOIDAL Q-ENRICHED CATEGORY DEFINITION        %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{definition}[Monoidal $\mathbb{Q}$-enriched category]
-$(E, \otimes)$ is strictly monoidal if
-\[
-\catE(a, b) \oplus \catE(a', b') \succeq \catE(a \otimes a', b \otimes b')
-\]
-and symmetric if
-\[
-e \succeq \catE(a \otimes b, b \otimes a)
-\]
-\end{definition}
 
@@
 \end{corollary}
+
+\section{Quantale Enriched Categories}
+\label{sec:quant-enrich-categ}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%        QUANTALE DEFINITION        %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{definition}[Quantale $\mathbb{Q}$]
+A quantale is a \emph{lattice} with a binary operator such that for all elements of the lattice $x, y$ and $\{x_i\}_i, \{y_i\}_i$
+\begin{align}
+x \oplus \bigvee y_i &= \bigvee x \oplus y_i \cr
+\left(\bigvee x_i\right) \oplus y &= \bigvee x_i \oplus y
+\end{align}
+
+We require that for all $x$, it exist $x^{-1}$ such that
+\begin{align}
+x \oplus x^{-1} &= e \cr
+x^{-1} \oplus x &= e
+\end{align}
+
+Such a quantale can be seen as a monoidal category over a \emph{group}.
+\end{definition}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%        Q-ENRICHED CATEGORY DEFINITION        %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{definition}[$\mathbb{Q}$-enriched category]
+A $\mathbb{Q}$-enriched category \catE is
+
+\begin{itemize}
+\item A set of objects
+\item For all objects $(a, b)$, $\catE(a, b) \in \mathbb{Q}$
+\end{itemize}
+
+such that for all object $a$
+\[
+e \succeq \catE(a, a)
+\]
+for all objects (a, b, c)
+\[
+\catE(b, c) \oplus \catE(a, b) \succeq \catE(a, c)
+\]
+\end{definition}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%        MONOIDAL Q-ENRICHED CATEGORY DEFINITION        %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{definition}[Monoidal $\mathbb{Q}$-enriched category]
+$(E, \otimes)$ is strictly monoidal if
+\[
+\catE(a, b) \oplus \catE(a', b') \succeq \catE(a \otimes a', b \otimes b')
+\]
+and symmetric if
+\[
+e \succeq \catE(a \otimes b, b \otimes a)
+\]
+\end{definition}
 

### 2018-08-06 article.tex
80 columns

@@
 Since $\circ$ and $\otimes$ are bifunctorial, we have the following
-interchange equations for all 0-cells $A, B \in \catC$ and all 2-cells $r, s, t, u \in \catC(A, B)$
+interchange equations for all 0-cells $A, B \in \catC$ and all 2-cells $r, s,
+    t, u \in \catC(A, B)$
 \begin{align}
@@
 \item 0-cells are natural number $\mathbb{N}$
-\item 1-cells are quantum circuits freely generated by $\circ$, $\otimes$ and $\{f_i\}_i$, a set of quantum circuits
-\item 2-cells are quantum circuit transformations freely generated by $\circ$, $\otimes$ and $\{r_i\}_i$, a set of ransformations, such that $\forall r: f \Rightarrow g \in \{r_i\}_i, \exists r^{-1}: g \Rightarrow f \in \{r_i\}_i$
+\item 1-cells are quantum circuits freely generated by $\circ$, $\otimes$ and
+    $\{f_i\}_i$, a set of quantum circuits
+\item 2-cells are quantum circuit transformations freely generated by $\circ$,
+    $\otimes$ and $\{r_i\}_i$, a set of ransformations, such that $\forall r: f
+    \Rightarrow g \in \{r_i\}_i, \exists r^{-1}: g \Rightarrow f \in \{r_i\}_i$
 \[
@@
 \item on 0-cell, $a \otimes b = a + b$
-\item on $\catC(a, b)$, a functor $\catC(a, b) \otimes C(a', b') \to C(a + a', b + b')$
+\item on $\catC(a, b)$, a functor $\catC(a, b) \otimes C(a', b') \to C(a + a',
+    b + b')$
 \end{itemize}
@@
 \begin{proof}
-We'll prove it by induction of the size of the transformation using the properties of the bifunctors.
+We'll prove it by induction of the size of the transformation using the
+    properties of the bifunctors.
 With a single transformation, we already have
@@
 \begin{align}
-r &= r_1 \bullet r_2 \bullet r_3 \bullet \cdots \bullet (s_1 \circ s_3) \bullet (s_2 \circ s_4) \cr
+r &= r_1 \bullet r_2 \bullet r_3 \bullet \cdots \bullet (s_1 \circ s_3) \bullet
+    (s_2 \circ s_4) \cr
 r &= r_1 \bullet r_2 \bullet r_3 \bullet \cdots \bullet r_{i - 1} \bullet r_i
 \end{align}
-Otherwise we can always add an identity transformation to get the interchange law.
+Otherwise we can always add an identity transformation to get the interchange
+    law.
 \begin{align}
-s_1 \circ (s_3 \bullet s_4) &\Rightarrow (s_1 \bullet id) \circ (s_2 \bullet s_3) \cr
-(s_1 \bullet s_2) \circ s_3 &\Rightarrow (s_1 \bullet s_2) \circ (id \bullet s_3)
+s_1 \circ (s_3 \bullet s_4) &\Rightarrow (s_1 \bullet id) \circ (s_2 \bullet
+    s_3) \cr
+(s_1 \bullet s_2) \circ s_3 &\Rightarrow (s_1 \bullet s_2) \circ (id \bullet
+    s_3)
 \end{align}
@@
 \begin{definition}[Quantale $\mathbb{Q}$]
-A quantale is a \emph{lattice} with a binary operator such that for all elements of the lattice $x, y$ and $\{x_i\}_i, \{y_i\}_i$
+A quantale is a \emph{lattice} with a binary operator such that for all
+    elements of the lattice $x, y$ and $\{x_i\}_i, \{y_i\}_i$
 \begin{align}
@@
 \begin{definition}[$\catC(a, b)$ to $\mathbb{Q}$ homomorphism]
-Let's define the following homomorphism $F$ from $\catC(a, b)$ to $\mathbb{Q}$. For all $f: a \to b \in \{f_i\}_i$
+Let's define the following homomorphism $F$ from $\catC(a, b)$ to $\mathbb{Q}$.
+    For all $f: a \to b \in \{f_i\}_i$
 \[
@@
 \begin{proof}
-The set of all possible transformation is freely generated by $\{r_i\}_i$, $\bullet$, $\circ$ and $\otimes$ so both $r$ and $s$ can be written
+The set of all possible transformation is freely generated by $\{r_i\}_i$,
+    $\bullet$, $\circ$ and $\otimes$ so both $r$ and $s$ can be written
 \begin{align}
@@
 &= V(r_1) \oplus V(r_2) \oplus \cdots \oplus V(r_n) \cr
-&= V(f \Rightarrow h_1) \oplus V(h_1 \Rightarrow h_2) \oplus \cdots \oplus V(h_n \Rightarrow g) \cr
-&= V(f) \oplus V(h_1)^{-1} \oplus V(h_1) \oplus V(h_2)^{-1} \oplus \cdots \oplus V(h_n) \oplus V(g)^{-1}
+&= V(f \Rightarrow h_1) \oplus V(h_1 \Rightarrow h_2) \oplus \cdots \oplus
+    V(h_n \Rightarrow g) \cr
+&= V(f) \oplus V(h_1)^{-1} \oplus V(h_1) \oplus V(h_2)^{-1} \oplus \cdots
+    \oplus V(h_n) \oplus V(g)^{-1}
 \end{align}
@@
 \[
-V(f) \oplus V(h_1)^{-1} \oplus V(h_1) \oplus V(h_2)^{-1} \oplus V(h_2) \oplus \cdots \oplus V(h_n) \oplus V(g)^{-1} = V(f) \oplus V(g)^{-1}
+V(f) \oplus V(h_1)^{-1} \oplus V(h_1) \oplus V(h_2)^{-1} \oplus V(h_2) \oplus
+    \cdots \oplus V(h_n) \oplus V(g)^{-1} = V(f) \oplus V(g)^{-1}
 \]
@@
 \begin{corollary}
-For all transformations $r: f \Rightarrow g$, then $V(r) = V(f) \oplus V(g)^{-1}$.
+For all transformations $r: f \Rightarrow g$, then $V(r) = V(f) \oplus
+    V(g)^{-1}$.
 Note that the definition of $F$ gives that only for $r \in \{f_i\}_i$.
@@
 \[
-\catE(f, g) = \begin{cases} {\rm if~} f \Rightarrow g \in \catC, &V(f \Rightarrow g) \cr {\rm otherwise~} &\top \end{cases}
+\catE(f, g) = \begin{cases} {\rm if~} f \Rightarrow g \in \catC, &V(f
+    \Rightarrow g) \cr {\rm otherwise~} &\top \end{cases}
 \]
@@
 \begin{proof}
-For all 1-cells $(f, g)$, if it exists a unique transformation $f \Rightarrow g$, then
+For all 1-cells $(f, g)$, if it exists a unique transformation $f \Rightarrow
+    g$, then
 \begin{align}
@@
 otherwise we already have $\catE(f, g) = \top \in \mathbb{Q}$.
-For all 1-cell $f$, we know that there is a identity transformation $f \Rightarrow f$ and
+For all 1-cell $f$, we know that there is a identity transformation $f
+    \Rightarrow f$ and
 \begin{align}
@@
 \end{align}
-Finally, for all 1-cells $(f, g, h)$ such that there exists $r: f \Rightarrow g$ and $s: g \Rightarrow h$ in $\catC$, then the homomorphism $F$ give us
+Finally, for all 1-cells $(f, g, h)$ such that there exists $r: f \Rightarrow
+    g$ and $s: g \Rightarrow h$ in $\catC$, then the homomorphism $F$ give us
 \begin{align}
@@
 \begin{proposition}
-$\catC(\mathbb{Q})$ is strictly symmetric monoidal with $\otimes$, the 2-functor of $\catC$.
+$\catC(\mathbb{Q})$ is strictly symmetric monoidal with $\otimes$, the
+    2-functor of $\catC$.
 
@@
 \[
-V(r: f \Rightarrow g) \oplus V(s: f' \Rightarrow g') = V(r \otimes s: f \otimes f' \Rightarrow g \otimes g')
+V(r: f \Rightarrow g) \oplus V(s: f' \Rightarrow g') = V(r \otimes s: f \otimes
+    f' \Rightarrow g \otimes g')
 \]
@@
 \]
-For the symmetry, we use the fact that $V(f \Rightarrow g) \oplus V(g \Rightarrow f) = e$
+For the symmetry, we use the fact that $V(f \Rightarrow g) \oplus V(g
+    \Rightarrow f) = e$
 \begin{align}
@@
 \begin{corollary}
-Given $s: f \Rightarrow g$ a transformation, and $s^{-1}: g \Rightarrow f$ the inverse transformation
+Given $s: f \Rightarrow g$ a transformation, and $s^{-1}: g \Rightarrow f$ the
+    inverse transformation
 \[

### 2018-08-06 article.tex
tab width 2
Added Q-circuits example at beginning

@@
 \input{preamble.tex}
-%%% some diagrams we won;t need in this papaer
-%\input{tikzpreamble.tex}
-%\input{quantumcircuits.tex}
+\input{tikzpreamble.tex}
+\input{quantumcircuits.tex}
 
@@
 \begin{example}  \label{ex:quantum-circs}
-  TODO -- quanutm circuits
+  The main example of interest is \Circ, the PROP of quantum circuits.
+  Different authors, and different hardware manufacturers, prefer
+  different primitive gate sets but the following is typical:
+  \[
+  \begin{array}{cccccccccc}
+    \circZ{{}} & \qquad  & \circX{{}} & \qquad & \circH \\ 
+    \vstrut{1.2em}
+    Z : 1 \to 1 && X : 1 \to 1 && H : 1 \to 1\\
+    \\
+    \circbox{R_Z(\theta)} & \qquad & \circCPhase{\theta} & \qquad & \circCX  \\ 
+    \vstrut{1.2em}
+    R_Z(\theta) : 1 \to 1 &&  \CPhase\theta :2 \to 2 && \CX : 2 \to 2\\
+  \end{array}
+  \]
+  We omit the usual matrix interpretations but these can be found in
+  the standard textbook \cite{NieChu:QuantComp:2000}.  Note that in
+  this example we actually have an uncountable family of generating
+  gates, one for each value $\theta \in [0,2\pi)$.  Unlike the case of
+  Boolean circuits all quantum gates have the same number of inputs and
+  outputs.  In fact, \Circ is a \dag-PROP, but this will not concern
+  us here.
 \end{example}

### 2018-08-06 article.tex

@@
 \begin{definition}  \label{def:strict-two-cat}
-A \emph{strict} 2-category \catC is a category \emph{enriched over
+A \emph{strict} 2-category $\mathcal{C}$ is a category \emph{enriched over
   \textbf{Cat}}, the category of categories, \emph{viz:}
 \begin{itemize}
-\item A category \catC, whose objects and arrows are known as
+\item A category $\mathcal{C}$, whose objects and arrows are known as
   0-cells and 1-cells respectively.
 \item For every pair of objects $A$ and $B$, we have a
-  category $\catC(A, B)$, whose arrows are known as 2-cells.
+  category $\mathcal{C}(A, B)$, whose arrows are known as 2-cells.
 \item For 2-cells $\alpha : f \Rightarrow g$ and $\beta : g
-  \Rightarrow h$ in $\catC(A,B)$ their (vertical) composite is denoted
+  \Rightarrow h$ in $\mathcal{C}(A,B)$ their (vertical) composite is denoted
   $\beta\bullet\alpha : f \Rightarrow h$.
@@
   $Id_f:f \Rightarrow f$.
-\item For all 0-cells $a$, $b$, $c$ a bifunctor $\circ: \catC(b, c)
-  \times \catC(a, b) \to \catC(a, c)$ called \emph{horizontal
+\item For all 0-cells $a$, $b$, $c$ a bifunctor $\circ: \mathcal{C}(b, c)
+  \times \mathcal{C}(a, b) \to \mathcal{C}(a, c)$ called \emph{horizontal
     composition} of 2-cells.
@@
 \begin{definition}\label{def:monoidal-two-cat}
-A \emph{strict} monoidal 2-category $(\catC, \otimes)$ is a strict
+A \emph{strict} monoidal 2-category $(\mathcal{C}, \otimes)$ is a strict
 2-category with
@@
 \begin{itemize}
-\item a 2-bifunctor $\otimes: \catC \times \catC \to \catC$ such that
-  on 2-cells $r, s, t, u$
+\item a 2-bifunctor $\otimes: \mathcal{C} \times \mathcal{C} \to \mathcal{C}$
+  such that on 2-cells $\alpha, \beta, \gamma, \delta$
 \[
-(r \otimes s) \circ (t \otimes u) = (r \circ t) \otimes (s \circ u)
+(\alpha \otimes \beta) \circ (\gamma \otimes \delta) = (\alpha \circ \gamma)
+  \otimes (\beta \circ \delta)
 \]
@@
 Since $\circ$ and $\otimes$ are bifunctorial, we have the following
-interchange equations for all 0-cells $A, B \in \catC$ and all 2-cells $r, s,
-    t, u \in \catC(A, B)$
+interchange equations for all 0-cells $A, B \in \mathcal{C}$ and all 2-cells
+  $\alpha, \beta, \gamma, \delta \in \mathcal{C}(A, B)$
 \begin{align}
-(r \circ s) \bullet (t \circ u) &= (r \bullet t) \circ (s \bullet u) \\
-(r \otimes s) \bullet (t \otimes u) &= (r \bullet t) \otimes (s \bullet u)
+(\alpha \circ \beta) \bullet (\gamma \circ \delta) &= (\alpha \bullet \gamma)
+  \circ (\beta \bullet \delta) \\
+(\alpha \otimes \beta) \bullet (\gamma \otimes \delta) &= (\alpha \bullet
+  \gamma) \otimes (\beta \bullet \delta)
 \end{align}
@@
 \begin{definition}
-A \emph{groupoid} is a category where morphisms are invertible.
-Let's define the following strict monoidal 2-groupoid \catC
-
+Given a free PROP of quantum circuits, we can extend it to a 2-PROP
+  $\mathcal{P}$ where
 \begin{itemize}
-\item 0-cells are natural number $\mathbb{N}$
-\item 1-cells are quantum circuits freely generated by $\circ$, $\otimes$ and
-  $\{f_i\}_i$, a set of quantum circuits
-\item 2-cells are quantum circuit transformations freely generated by $\circ$,
-  $\otimes$ and $\{r_i\}_i$, a set of ransformations, such that $\forall r: f
-    \Rightarrow g \in \{r_i\}_i, \exists r^{-1}: g \Rightarrow f \in \{r_i\}_i$
-\[
-r \circ r^{-1} = id_f
-\]
-\item the vertical composition is the usual transformations composition
-\item the horizontal composition is the usual circuits composition
-\item the 2-bifunctor is
+\item gates form the set of 1-cells 
+\item the 2-cells are generated by a finite set of \emph{rewriting} rules such
+  that
 \begin{itemize}
-\item on 0-cell, $a \otimes b = a + b$
-\item on $\catC(a, b)$, a functor $\catC(a, b) \otimes C(a', b') \to C(a + a',
-  b + b')$
+\item 2 rewriting rules with the same types must be equal
+\item each rewriting rule must be invertible, that is $\mathcal{C}(A, B)$ is a
+  \emph{groupoid} for all 0-cells $A, B$
 \end{itemize}
 \end{itemize}
+
+2-cells are in general called \emph{rewrite sequence}, but if a rewrite sequence
+  has only 1 occurence of a rewrite rule, it is called an \emph{atomic rule}.
 \end{definition}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%        REWRITING SYSTEM EXAMPLE        %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{example}
+TODO rewrite sequence \& atomic rule
+\end{example}
 
@@
 \begin{proposition}
-For all transformation $r: f \Rightarrow g$, $r$ can be writen
+For all rewriting sequences $\alpha: f \Rightarrow g$, $\alpha$ can be writen
 \[
-r = r_1 \bullet r_2 \bullet r_3 \bullet \cdots \bullet r_i
+\alpha = \alpha_1 \bullet \alpha_2 \bullet \alpha_3 \bullet \cdots \bullet
+  \alpha_i
 \]
@@
 \begin{proof}
-We'll prove it by induction of the size of the transformation using the
+We'll prove it by induction of the size of the rewriting sequence using the
   properties of the bifunctors.
-With a single transformation, we already have
+With a atomic rule, we already have
 \[
-r = r_1
+\alpha = \alpha_1
 \]
-For the inductive step let's assume $r$ is
+For the inductive step let's assume $\beta$ is
 \[
-r = r_1 \bullet r_2 \bullet r_3 \bullet \cdots \bullet s
+\alpha = \alpha_1 \bullet \alpha_2 \bullet \alpha_3 \bullet \cdots \bullet \beta
 \]
@@
 \[
-s = (s_1 \bullet s_2) \circ (s_3 \bullet s_4)
+\beta = (\beta_1 \bullet \beta_2) \circ (\beta_3 \bullet \beta_4)
 \]
@@
 \[
-s = (s_1 \circ s_3) \bullet (s_2 \circ s_4)
+\beta = (\beta_1 \circ \beta_3) \bullet (\beta_2 \circ \beta_4)
 \]
@@
 \begin{align}
-r &= r_1 \bullet r_2 \bullet r_3 \bullet \cdots \bullet (s_1 \circ s_3) \bullet
-  (s_2 \circ s_4) \cr
-r &= r_1 \bullet r_2 \bullet r_3 \bullet \cdots \bullet r_{i - 1} \bullet r_i
+\alpha &= \alpha_1 \bullet \alpha_2 \bullet \alpha_3 \bullet \cdots \bullet
+  (\beta_1 \circ \beta_3) \bullet (\beta_2 \circ \beta_4) \cr
+\alpha &= \alpha_1 \bullet \alpha_2 \bullet \alpha_3 \bullet \cdots \bullet
+  \alpha_{i - 1} \bullet \alpha_i
 \end{align}
-Otherwise we can always add an identity transformation to get the interchange
+Otherwise we can always add an identity rewriting rule to get the interchange
   law.
 \begin{align}
-s_1 \circ (s_3 \bullet s_4) &\Rightarrow (s_1 \bullet id) \circ (s_2 \bullet
-  s_3) \cr
-(s_1 \bullet s_2) \circ s_3 &\Rightarrow (s_1 \bullet s_2) \circ (id \bullet
-  s_3)
+\beta_1 \circ (\beta_3 \bullet \beta_4) &\Rightarrow (\beta_1 \bullet id) \circ
+  (\beta_2 \bullet \beta_3) \cr
+(\beta_1 \bullet \beta_2) \circ \beta_3 &\Rightarrow (\beta_1 \bullet \beta_2)
+  \circ (id \bullet \beta_3)
 \end{align}
@@
 \[
-s = (s_1 \bullet s_2) \otimes (s_3 \bullet s_4)
+\beta = (\beta_1 \bullet \beta_2) \otimes (\beta_3 \bullet \beta_4)
 \]
@@
 \begin{proposition}
-For all transformation $r: f \Rightarrow g$, $r$ can be writen
+For all rewriting sequences $\alpha: f \Rightarrow g$, $\alpha$ can be writen
 \begin{align}
-r &= r_1 \circ r_2 \circ r_3 \circ \cdots \circ r_j \cr
-r &= r_1 \otimes r_2 \otimes r_3 \otimes \cdots \otimes r_k
+\alpha &= \alpha_1 \circ \alpha_2 \circ \alpha_3 \circ \cdots \circ \alpha_j \cr
+\alpha &= \alpha_1 \otimes \alpha_2 \otimes \alpha_3 \otimes \cdots \otimes
+  \alpha_k
 \end{align}
@@
 \[
-(r \otimes s) \circ (t \otimes u) = (r \circ t) \otimes (s \circ u)
+(\alpha \otimes \beta) \circ (\gamma \otimes \delta) = (\alpha \circ \gamma)
+  \otimes (\beta \circ \delta)
 \]
@@
 \begin{corollary}
-For all transformation $r: f \Rightarrow g$, $r$ can be writen
+For all rewriting sequences $r: f \Rightarrow g$, $r$ can be writen
 \[
-r = r_1 \bullet r_2 \bullet r_3 \bullet \cdots \bullet r_i
+\alpha = \alpha_1 \bullet \alpha_2 \bullet \alpha_3 \bullet \cdots \bullet
+  \alpha_i
 \]
@@
 \[
-r_i = r_{i, 1} \circ r_{i, 2} \circ r_{i, 3} \circ \cdots \circ r_{i, j}
+\alpha_i = \alpha_{i, 1} \circ \alpha_{i, 2} \circ \alpha_{i, 3} \circ \cdots
+  \circ \alpha_{i, j}
 \]
@@
 \[
-r_{i, j} = \bigotimes_k^l r_{i, j, k}
+\alpha_{i, j} = \bigotimes_k^l \alpha_{i, j, k}
 \]
-a tensor products of transformations.
+a tensor products of rewriting sequences.
 \end{corollary}
@@
 \begin{definition}[$\mathbb{Q}$-enriched category]
-A $\mathbb{Q}$-enriched category \catE is
+A $\mathbb{Q}$-enriched category $\mathcal{E}$ is
 
@@
 \item A set of objects
-\item For all objects $(a, b)$, $\catE(a, b) \in \mathbb{Q}$
+\item For all objects $(a, b)$, $\mathcal{E}(a, b) \in \mathbb{Q}$
 \end{itemize}
 
-such that for all object $a$
+such that for all objects $a$
 \[
-e \succeq \catE(a, a)
+e \succeq \mathcal{E}(a, a)
 \]
@@
 \[
-\catE(b, c) \oplus \catE(a, b) \succeq \catE(a, c)
+\mathcal{E}(b, c) \oplus \mathcal{E}(a, b) \succeq \mathcal{E}(a, c)
 \]
@@
 \[
-\catE(a, b) \oplus \catE(a', b') \succeq \catE(a \otimes a', b \otimes b')
+\mathcal{E}(a, b) \oplus \mathcal{E}(a', b') \succeq \mathcal{E}(a \otimes a',
+  b \otimes b')
 \]
@@
 \[
-e \succeq \catE(a \otimes b, b \otimes a)
+e \succeq \mathcal{E}(a \otimes b, b \otimes a)
 \]
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{definition}[$\catC(a, b)$ to $\mathbb{Q}$ homomorphism]
-Let's define the following homomorphism $F$ from $\catC(a, b)$ to $\mathbb{Q}$.
-  For all $f: a \to b \in \{f_i\}_i$
+\begin{definition}
+For all $(A, B) \in \mathcal{P}$ we define a homomorphism $V: \mathcal{C}(A, B)
+  \to \mathbb{Q}$, such that for all gates $f$
 \[
@@
 \]
-For all $r: f \Rightarrow g \in \{r_i\}_i$
+For all rewriting rules $\alpha: f \Rightarrow g$
 \[
-V(r) = V(f) \oplus V(g)^{-1}
+V(\alpha) = V(f) \oplus V(g)^{-1}
 \]
 
-On 1-cells
+On gates
 \begin{align}
@@
 
-On 2-cells
+On rewriting sequences
 \begin{align}
-V(r \circ s) &= V(r) \oplus V(s) \cr
-V(r \otimes s) &= V(r) \oplus V(s) \cr
-V(r \bullet s) &= V(r) \oplus V(s)
+V(\alpha \circ \beta) &= V(\alpha) \oplus V(\beta) \cr
+V(\alpha \otimes \beta) &= V(\alpha) \oplus V(\beta) \cr
+V(\alpha \bullet \beta) &= V(\alpha) \oplus V(\beta)
 \end{align}
@@
 \[
-V(r: f \Rightarrow g) = V(f) \oplus V(g)^{-1}
+V(\alpha: f \Rightarrow g) = V(f) \oplus V(g)^{-1}
 \]
@@
 \begin{align}
-V(r: f \Rightarrow f) &= V(f) \oplus V(f)^{-1} \cr
+V(\alpha: f \Rightarrow f) &= V(f) \oplus V(f)^{-1} \cr
 &= e
@@
 \begin{proposition}
-Given two transformations $r, s: f \Rightarrow g$, then $V(r) = V(s)$.
+Given two rewriting sequence $\alpha, \beta: f \Rightarrow g$, then $V(\alpha) =
+  V(\beta)$.
 
 \begin{proof}
-The set of all possible transformation is freely generated by $\{r_i\}_i$,
-  $\bullet$, $\circ$ and $\otimes$ so both $r$ and $s$ can be written
+The set of all possible rewriting sequences is freely generated by the set of
+rewriting rules, $\bullet$, $\circ$ and $\otimes$ so both $\alpha$ and $\beta$
+  can be written
 \begin{align}
-r &= r_1 \odot r_2 \odot r_3 \odot \cdots \odot r_n \cr
-s &= s_1 \odot s_2 \odot s_3 \odot \cdots \odot s_m
+\alpha &= \alpha_1 \odot \alpha_2 \odot \alpha_3 \odot \cdots \odot \alpha_n \cr
+\beta &= \beta_1 \odot \beta_2 \odot \beta_3 \odot \cdots \odot \beta_m
 \end{align}
-with $\odot \in \{\bullet, \circ, \otimes\}$ and for all i
-\begin{align}
-r_i &\in \{f_i\}_i \cr
-s_i &\in \{f_i\}_i
-\end{align}
-If $n = 1$ then $r \in \{f_i\}_i$, so $V(r) = V(f) \oplus V(g)^{-1}$.
+with $\odot \in \{\bullet, \circ, \otimes\}$ and for all i $\alpha_i$ and
+$\beta_i$ are rewriting rule.
+If $n = 1$ then $\alpha$ is a rewriting rule, so $V(r) = V(f) \oplus V(g)^{-1}$.
 Otherwise
 \begin{align}
-V(r) &= V(r_1 \odot r_2 \odot \cdots \odot r_n) \cr
-&= V(r_1) \oplus V(r_2) \oplus \cdots \oplus V(r_n) \cr
+V(\alpha) &= V(\alpha_1 \odot \alpha_2 \odot \cdots \odot \alpha_n) \cr
+&= V(\alpha_1) \oplus V(\alpha_2) \oplus \cdots \oplus V(\alpha_n) \cr
 &= V(f \Rightarrow h_1) \oplus V(h_1 \Rightarrow h_2) \oplus \cdots \oplus
@@
 \]
-The same is true for $s$, so
+The same is true for $\beta$, so
 \[
-V(r) = V(s) = V(f) \oplus V(g)^{-1}
+V(\alpha) = V(\beta) = V(f) \oplus V(g)^{-1}
 \]
@@
 \begin{corollary}
-For all transformations $r: f \Rightarrow g$, then $V(r) = V(f) \oplus
-  V(g)^{-1}$.
-Note that the definition of $F$ gives that only for $r \in \{f_i\}_i$.
+For all rewriting sequences $\alpha: f \Rightarrow g$, then $V(\alpha) = V(f)
+  \oplus V(g)^{-1}$.
+Note that the definition of $V$ gives that only for rewriting rules.
 \end{corollary}
@@
 \begin{proposition}
-The following category $\catC(\mathbb{Q})$ is a $\mathbb{Q}$-category
+The following category $\mathcal{P}(\mathbb{Q})$ is a $\mathbb{Q}$-category
 \begin{itemize}
-\item A set of objects which are the 1-cells of $\catC$
+\item A set of objects which are the gates of $\mathcal{P}$
 \item For all objects $(f, g)$
@@
 \[
-\catE(f, g) = \begin{cases} {\rm if~} f \Rightarrow g \in \catC, &V(f
-  \Rightarrow g) \cr {\rm otherwise~} &\top \end{cases}
+\mathcal{E}(f, g) =
+\begin{cases}
+\textrm{if } f \Rightarrow g \in \mathcal{P}, &V(f \Rightarrow g) \cr
+\textrm{otherwise } &\top
+\end{cases}
 \]
-With $\top \in Q$ such that for all $x \in \mathbb{Q}$
+With $\top \in \mathbb{Q}$ such that for all $x \in \mathbb{Q}$
 \[
@@
 \begin{proof}
-For all 1-cells $(f, g)$, if it exists a unique transformation $f \Rightarrow
+For all gates $(f, g)$, if it exists a unique transformation $f \Rightarrow
   g$, then
 \begin{align}
-\catE(f, g) &= V(f \Rightarrow g) \in \mathbb{Q} \cr
+\mathcal{E}(f, g) &= V(f \Rightarrow g) \in \mathbb{Q} \cr
 &= V(f) \oplus V(g)^{-1}
 \end{align}
-otherwise we already have $\catE(f, g) = \top \in \mathbb{Q}$.
-For all 1-cell $f$, we know that there is a identity transformation $f
+otherwise we already have $\mathcal{E}(f, g) = \top \in \mathbb{Q}$.
+For all gates $f$, we know that there is a identity transformation $f
   \Rightarrow f$ and
 \begin{align}
-\catE(f, f) &= V(id_f) \cr
+\mathcal{E}(f, f) &= V(id_f) \cr
 &= V(f \Rightarrow f)\cr
@@
 \end{align}
-Finally, for all 1-cells $(f, g, h)$ such that there exists $r: f \Rightarrow
-  g$ and $s: g \Rightarrow h$ in $\catC$, then the homomorphism $F$ give us
+Finally, for all gates $(f, g, h)$ such that there exists $\alpha: f \Rightarrow
+  g$ and $\beta: g \Rightarrow h$ in $\mathcal{P}$, then the homomorphism $V$
+    give us
 \begin{align}
-\catE(f, h) &= V(f \Rightarrow h) \cr
-&= V(s\circ r) \cr
-&= V(s) \oplus V(r) \cr
+\mathcal{E}(f, h) &= V(f \Rightarrow h) \cr
+&= V(\alpha \circ \beta) \cr
+&= V(\alpha) \oplus V(\beta) \cr
 &= V(g \Rightarrow h) \oplus V(f \Rightarrow g) \cr
-&= \catE(g, h) \oplus \catE(f, g)
+&= \mathcal{E}(g, h) \oplus \mathcal{E}(f, g)
 \end{align}
-If either $r$ or $s$ doesn't exists in $\catC$ then
+If either $\alpha$ or $\beta$ doesn't exists in $\mathcal{P}$ then
 \[
-\catE(b, c) \oplus \catE(a, b) \succeq \catE(a, c)
+\mathcal{E}(b, c) \oplus \mathcal{E}(a, b) \succeq \mathcal{E}(a, c)
 \]
@@
 \[\begin{matrix}
-&\catE(b, c) \oplus \top &\succeq \top \cr
-{\rm or~} &\top \oplus \catE(a, b) &\succeq \top \cr
-{\rm or~} &\top \oplus \top &\succeq \top
+&\mathcal{E}(b, c) \oplus \top &\succeq \top \cr
+\textrm{or} &\top \oplus \mathcal{E}(a, b) &\succeq \top \cr
+\textrm{or} &\top \oplus \top &\succeq \top
 \end{matrix}\]
@@
 \begin{proposition}
-$\catC(\mathbb{Q})$ is strictly symmetric monoidal with $\otimes$, the
-  2-functor of $\catC$.
+$\mathcal{P}(\mathbb{Q})$ is strictly symmetric monoidal with $\otimes$, the
+  2-functor of $\mathcal{P}$.
 
@@
 \[
-V(r) \oplus V(s) \succeq V(r \otimes s)
+V(\alpha) \oplus V(\beta) \succeq V(\alpha \otimes \beta)
 \]
@@
 \[
-V(r: f \Rightarrow g) \oplus V(s: f' \Rightarrow g') = V(r \otimes s: f \otimes
-  f' \Rightarrow g \otimes g')
+V(\alpha: f \Rightarrow g) \oplus V(\beta: f' \Rightarrow g') = V(\alpha \otimes
+  \beta: f \otimes f' \Rightarrow g \otimes g')
 \]
@@
 \[
-\catE(f, g) \oplus \catE(f', g') = \catE(f \otimes f', g \otimes g')
+\mathcal{E}(f, g) \oplus \mathcal{E}(f', g') = \mathcal{E}(f \otimes f', g
+  \otimes g')
 \]
@@
 \begin{align}
-\catE(f \otimes g, g \otimes f) &= V(f \otimes g \Rightarrow g \otimes g) \cr
+\mathcal{E}(f \otimes g, g \otimes f) &= V(f \otimes g \Rightarrow g \otimes g)
+  \cr
 &= V(f \Rightarrow g \otimes g \Rightarrow f) \cr
@@
 \begin{corollary}
-Given $s: f \Rightarrow g$ a transformation, and $s^{-1}: g \Rightarrow f$ the
-  inverse transformation
+Given $\alpha: f \Rightarrow g$ a rewriting sequence, and $\alpha^{-1}: g
+  \Rightarrow f$ the inverse rewriting sequence
 \[

### 2018-08-06 article.tex
typo and r, s, t, u to alpha, beta, gamma, delta

@@
 \input{preamble.tex}
-\input{tikzpreamble.tex}
-\input{quantumcircuits.tex}
+%%% some diagrams we won;t need in this papaer
+%\input{tikzpreamble.tex}
+%\input{quantumcircuits.tex}
 
@@
 \begin{example}  \label{ex:quantum-circs}
-  The main example of interest is \Circ, the PROP of quantum circuits.
-  Different authors, and different hardware manufacturers, prefer
-  different primitive gate sets but the following is typical:
-  \[
-  \begin{array}{cccccccccc}
-    \circZ{{}} & \qquad  & \circX{{}} & \qquad & \circH \\ 
-    \vstrut{1.2em}
-    Z : 1 \to 1 && X : 1 \to 1 && H : 1 \to 1\\
-    \\
-    \circbox{R_Z(\theta)} & \qquad & \circCPhase{\theta} & \qquad & \circCX  \\ 
-    \vstrut{1.2em}
-    R_Z(\theta) : 1 \to 1 &&  \CPhase\theta :2 \to 2 && \CX : 2 \to 2\\
-  \end{array}
-  \]
-  We omit the usual matrix interpretations but these can be found in
-  the standard textbook \cite{NieChu:QuantComp:2000}.  Note that in
-  this example we actually have an uncountable family of generating
-  gates, one for each value $\theta \in [0,2\pi)$.  Unlike the case of
-  Boolean circuits all quantum gates have the same number of inputs and
-  outputs.  In fact, \Circ is a \dag-PROP, but this will not concern
-  us here.
+  TODO -- quanutm circuits
 \end{example}
@@
 \begin{definition}  \label{def:strict-two-cat}
-A \emph{strict} 2-category $\mathcal{C}$ is a category \emph{enriched over
+A \emph{strict} 2-category \catC is a category \emph{enriched over
   \textbf{Cat}}, the category of categories, \emph{viz:}
 \begin{itemize}
-\item A category $\mathcal{C}$, whose objects and arrows are known as
+\item A category \catC, whose objects and arrows are known as
   0-cells and 1-cells respectively.
 \item For every pair of objects $A$ and $B$, we have a
-  category $\mathcal{C}(A, B)$, whose arrows are known as 2-cells.
+  category $\catC(A, B)$, whose arrows are known as 2-cells.
 \item For 2-cells $\alpha : f \Rightarrow g$ and $\beta : g
-  \Rightarrow h$ in $\mathcal{C}(A,B)$ their (vertical) composite is denoted
+  \Rightarrow h$ in $\catC(A,B)$ their (vertical) composite is denoted
   $\beta\bullet\alpha : f \Rightarrow h$.
@@
   $Id_f:f \Rightarrow f$.
-\item For all 0-cells $a$, $b$, $c$ a bifunctor $\circ: \mathcal{C}(b, c)
-  \times \mathcal{C}(a, b) \to \mathcal{C}(a, c)$ called \emph{horizontal
+\item For all 0-cells $a$, $b$, $c$ a bifunctor $\circ: \catC(b, c)
+  \times \catC(a, b) \to \catC(a, c)$ called \emph{horizontal
     composition} of 2-cells.
@@
 \begin{definition}\label{def:monoidal-two-cat}
-A \emph{strict} monoidal 2-category $(\mathcal{C}, \otimes)$ is a strict
+A \emph{strict} monoidal 2-category $(\catC, \otimes)$ is a strict
 2-category with
@@
 \begin{itemize}
-\item a 2-bifunctor $\otimes: \mathcal{C} \times \mathcal{C} \to \mathcal{C}$
-  such that on 2-cells $\alpha, \beta, \gamma, \delta$
+\item a 2-bifunctor $\otimes: \catC \times \catC \to \catC$ such that
+  on 2-cells $\alpha, \beta, \gamma, \delta$
 \[
@@
 Since $\circ$ and $\otimes$ are bifunctorial, we have the following
-interchange equations for all 0-cells $A, B \in \mathcal{C}$ and all 2-cells
-  $\alpha, \beta, \gamma, \delta \in \mathcal{C}(A, B)$
+interchange equations for all 0-cells $A, B \in \catC$ and all 2-cells $\alpha,
+  \beta, \gamma, \delta \in \catC(A, B)$
 \begin{align}
@@
 \begin{definition}
-Given a free PROP of quantum circuits, we can extend it to a 2-PROP
-  $\mathcal{P}$ where
+Given a free PROP of quantum circuits, we can extend it to a 2-PROP $\mathbb{P}$
+  where
 \begin{itemize}
@@
 \item 2 rewriting rules with the same types must be equal
-\item each rewriting rule must be invertible, that is $\mathcal{C}(A, B)$ is a
+\item each rewriting rule must be invertible, that is $\catC(A, B)$ is a
   \emph{groupoid} for all 0-cells $A, B$
@@
 \[
-\alpha = \alpha_1 \bullet \alpha_2 \bullet \alpha_3 \bullet \cdots \bullet
-  \alpha_i
+\alpha = \alpha_1 \bullet \alpha_2 \bullet \alpha_3 \bullet \cdots \bullet \alpha_i
 \]
@@
 \begin{definition}[$\mathbb{Q}$-enriched category]
-A $\mathbb{Q}$-enriched category $\mathcal{E}$ is
+A $\mathbb{Q}$-enriched category \catE is
 
@@
 \item A set of objects
-\item For all objects $(a, b)$, $\mathcal{E}(a, b) \in \mathbb{Q}$
+\item For all objects $(a, b)$, $\catE(a, b) \in \mathbb{Q}$
 \end{itemize}
@@
 \[
-e \succeq \mathcal{E}(a, a)
+e \succeq \catE(a, a)
 \]
@@
 \[
-\mathcal{E}(b, c) \oplus \mathcal{E}(a, b) \succeq \mathcal{E}(a, c)
+\catE(b, c) \oplus \catE(a, b) \succeq \catE(a, c)
 \]
@@
 \[
-\mathcal{E}(a, b) \oplus \mathcal{E}(a', b') \succeq \mathcal{E}(a \otimes a',
-  b \otimes b')
+\catE(a, b) \oplus \catE(a', b') \succeq \catE(a \otimes a', b \otimes b')
 \]
@@
 \[
-e \succeq \mathcal{E}(a \otimes b, b \otimes a)
+e \succeq \catE(a \otimes b, b \otimes a)
 \]
@@
 \begin{definition}
-For all $(A, B) \in \mathcal{P}$ we define a homomorphism $V: \mathcal{C}(A, B)
-  \to \mathbb{Q}$, such that for all gates $f$
+For all $(A, B) \in \mathbb{P}$ we define a homomorphism $V: \catC(A, B) \to
+  \mathbb{Q}$, such that for all gates $f$
 \[
@@
 \begin{proposition}
-The following category $\mathcal{P}(\mathbb{Q})$ is a $\mathbb{Q}$-category
+The following category $\mathbb{P}(\mathbb{Q})$ is a $\mathbb{Q}$-category
 \begin{itemize}
-\item A set of objects which are the gates of $\mathcal{P}$
+\item A set of objects which are the gates of $\mathbb{P}$
 \item For all objects $(f, g)$
@@
 \[
-\mathcal{E}(f, g) =
+\catE(f, g) =
 \begin{cases}
-\textrm{if } f \Rightarrow g \in \mathcal{P}, &V(f \Rightarrow g) \cr
-\textrm{otherwise } &\top
+{\rm if~} f \Rightarrow g \in \mathbb{P}, &V(f \Rightarrow g) \cr
+{\rm otherwise~} &\top
 \end{cases}
@@
 \begin{align}
-\mathcal{E}(f, g) &= V(f \Rightarrow g) \in \mathbb{Q} \cr
+\catE(f, g) &= V(f \Rightarrow g) \in \mathbb{Q} \cr
 &= V(f) \oplus V(g)^{-1}
 \end{align}
-otherwise we already have $\mathcal{E}(f, g) = \top \in \mathbb{Q}$.
+otherwise we already have $\catE(f, g) = \top \in \mathbb{Q}$.
 For all gates $f$, we know that there is a identity transformation $f
@@
 \begin{align}
-\mathcal{E}(f, f) &= V(id_f) \cr
+\catE(f, f) &= V(id_f) \cr
 &= V(f \Rightarrow f)\cr
@@
 \end{align}
-Finally, for all gates $(f, g, h)$ such that there exists $\alpha: f \Rightarrow
-  g$ and $\beta: g \Rightarrow h$ in $\mathcal{P}$, then the homomorphism $V$
-    give us
+Finally, for all gates $(f, g, h)$ such that there exists $r: f \Rightarrow
+  g$ and $s: g \Rightarrow h$ in $\catC$, then the homomorphism $V$ give us
 \begin{align}
-\mathcal{E}(f, h) &= V(f \Rightarrow h) \cr
+\catE(f, h) &= V(f \Rightarrow h) \cr
 &= V(\alpha \circ \beta) \cr
-&= V(\alpha) \oplus V(\beta) \cr
+&= V(\align) \oplus V(\beta) \cr
 &= V(g \Rightarrow h) \oplus V(f \Rightarrow g) \cr
-&= \mathcal{E}(g, h) \oplus \mathcal{E}(f, g)
+&= \catE(g, h) \oplus \catE(f, g)
 \end{align}
-If either $\alpha$ or $\beta$ doesn't exists in $\mathcal{P}$ then
+If either $\alpha$ or $\beta$ doesn't exists in $\mathbb{P}$ then
 \[
-\mathcal{E}(b, c) \oplus \mathcal{E}(a, b) \succeq \mathcal{E}(a, c)
+\catE(b, c) \oplus \catE(a, b) \succeq \catE(a, c)
 \]
@@
 \[\begin{matrix}
-&\mathcal{E}(b, c) \oplus \top &\succeq \top \cr
-\textrm{or} &\top \oplus \mathcal{E}(a, b) &\succeq \top \cr
-\textrm{or} &\top \oplus \top &\succeq \top
+&\catE(b, c) \oplus \top &\succeq \top \cr
+{\rm or~} &\top \oplus \catE(a, b) &\succeq \top \cr
+{\rm or~} &\top \oplus \top &\succeq \top
 \end{matrix}\]
@@
 \begin{proposition}
-$\mathcal{P}(\mathbb{Q})$ is strictly symmetric monoidal with $\otimes$, the
-  2-functor of $\mathcal{P}$.
+$\mathbb{P}(\mathbb{Q})$ is strictly symmetric monoidal with $\otimes$, the
+  2-functor of $\mathbb{P}$.
 
@@
 \[
-\mathcal{E}(f, g) \oplus \mathcal{E}(f', g') = \mathcal{E}(f \otimes f', g
-  \otimes g')
+\catE(f, g) \oplus \catE(f', g') = \catE(f \otimes f', g \otimes g')
 \]
@@
 \begin{align}
-\mathcal{E}(f \otimes g, g \otimes f) &= V(f \otimes g \Rightarrow g \otimes g)
-  \cr
+\catE(f \otimes g, g \otimes f) &= V(f \otimes g \Rightarrow g \otimes g) \cr
 &= V(f \Rightarrow g \otimes g \Rightarrow f) \cr

### 2018-08-06 article.tex
LaTeX mistakes

@@
 \end{align}
-Finally, for all gates $(f, g, h)$ such that there exists $r: f \Rightarrow
-  g$ and $s: g \Rightarrow h$ in $\catC$, then the homomorphism $V$ give us
+Finally, for all gates $(f, g, h)$ such that there exists $\alpha: f \Rightarrow
+  g$ and $\beta: g \Rightarrow h$ in $\mathbb{P}$, then the homomorphism $V$
+    give us
 \begin{align}
@@
 &= V(\alpha \circ \beta) \cr
-&= V(\align) \oplus V(\beta) \cr
+&= V(\alpha) \oplus V(\beta) \cr
 &= V(g \Rightarrow h) \oplus V(f \Rightarrow g) \cr

### 2018-08-06 article.tex
2-functor definition and format consistency

@@
 
-\begin{definition}\label{def:PROP}
+\begin{definition} \label{def:PROP}
   A \emph{PROP} is strict symmetric monoidal category whose objects
@@
 \begin{example} \label{ex:quantum-circs}
-  TODO -- quanutm circuits
+  The main example of interest is \Circ, the PROP of quantum circuits.
+  Different authors, and different hardware manufacturers, prefer
+  different primitive gate sets but the following is typical:
+  \[
+  \begin{array}{cccccccccc}
+    \circZ{{}} & \qquad  & \circX{{}} & \qquad & \circH \\ 
+    \vstrut{1.2em}
+    Z : 1 \to 1 && X : 1 \to 1 && H : 1 \to 1\\
+    \\
+    \circbox{R_Z(\theta)} & \qquad & \circCPhase{\theta} & \qquad & \circCX  \\ 
+    \vstrut{1.2em}
+    R_Z(\theta) : 1 \to 1 &&  \CPhase\theta :2 \to 2 && \CX : 2 \to 2\\
+  \end{array}
+  \]
+  We omit the usual matrix interpretations but these can be found in
+  the standard textbook \cite{NieChu:QuantComp:2000}.  Note that in
+  this example we actually have an uncountable family of generating
+  gates, one for each value $\theta \in [0,2\pi)$.  Unlike the case of
+  Boolean circuits all quantum gates have the same number of inputs and
+  outputs.  In fact, \Circ is a \dag-PROP, but this will not concern
+  us here.
 \end{example}
@@
     \item A category $\mathcal{C}$, whose objects and arrows are known as
-  0-cells and 1-cells respectively.
+      \emph{0-cells} and \emph{1-cells} respectively.
     \item For every pair of objects $A$ and $B$, we have a
-  category $\mathcal{C}(A, B)$, whose arrows are known as 2-cells.
-\item For 2-cells $\alpha : f \Rightarrow g$ and $\beta : g
+      category $\mathcal{C}(A, B)$, whose arrows are known as \emph{2-cells}.
+    \item For \emph{2-cells} $\alpha : f \Rightarrow g$ and $\beta : g
       \Rightarrow h$ in $\mathcal{C}(A,B)$ their (vertical) composite is denoted
-  $\beta\bullet\alpha : f \Rightarrow h$.
+      $\beta \bullet \alpha : f \Rightarrow h$.
     \item For every 1-cell $f:A\to B$ we denote its identity 2-cell by
       $Id_f:f \Rightarrow f$.
-\item For all 0-cells $a$, $b$, $c$ a bifunctor $\circ: \mathcal{C}(b, c)
-  \times \mathcal{C}(a, b) \to \mathcal{C}(a, c)$ called \emph{horizontal
-    composition} of 2-cells.
+    \item For all \emph{0-cells} $a$, $b$, $c$ a bifunctor $\circ:
+      \mathcal{C}(b, c) \times \mathcal{C}(a, b) \to \mathcal{C}(a, c)$ called
+      \emph{horizontal composition} of \emph{2-cells}.
   \end{itemize}
-\REM{Need here the name for a 2-cat with all 2-cells invertible;  the
-  term 2-groupoid refers to the case where both 1- and 2-cells are
+\REM{Need here the name for a 2-cat with all \emph{2-cells} invertible;  the
+  term 2-groupoid refers to the case where both 1- and \emph{2-cells} are
   invertible.}
@@
 
-\begin{definition}
-  \label{def:two-functor}
-  DEFINITION OF 2-FUNCTOR here
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%        2-FUNCTOR DEFINITION        %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{definition} \label{def:two-functor}
+  A \emph{strict} 2-functor $F$ between two \emph{strict} 2-categories
+    $\mathcal{C}$ and $\mathcal{D}$ is
+  \begin{itemize}
+    \item for all \emph{0-cells} $A \in \mathcal{C}$ a 0-cell $F(A) \in
+      \mathcal{D}$
+    \item for all \emph{0-cells} $(A, B) \in \mathcal{C}$ a functor $F_{A, B}:
+      \mathcal{C}(A, B) \to \mathcal{D}(F(A), F(B))$
+  \end{itemize}
+
+  \noindent
+  Such that $F$ respect the horizontal and vertical compositions.
 \end{definition}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \begin{definition}\label{def:monoidal-two-cat}
   A \emph{strict} monoidal 2-category $(\mathcal{C}, \otimes)$ is a strict
-2-category with
-\REM{Is this \emph{required}?  Doesn't it follow from the  definition
-  of bifunctor?  (It does in the 1-category case, so it should do
-  here!)}
-\begin{itemize}
-\item a 2-bifunctor $\otimes: \mathcal{C} \times \mathcal{C} \to \mathcal{C}$
-  such that on 2-cells $\alpha, \beta, \gamma, \delta$
-\[
-(\alpha \otimes \beta) \circ (\gamma \otimes \delta) = (\alpha \circ \gamma)
-  \otimes (\beta \circ \delta)
-\]
+    2-category with a 2-bifunctor $\otimes: \mathcal{C} \times \mathcal{C} \to
+    \mathcal{C}$.
   \REM{This definition doesn't quite work without the additoin type
-  information on the 2-cells}
-\end{itemize}
+    information on the \emph{2-cells}}
 \end{definition}
@@
   Since $\circ$ and $\otimes$ are bifunctorial, we have the following
-interchange equations for all 0-cells $A, B \in \mathcal{C}$ and all 2-cells
-  $\alpha, \beta, \gamma, \delta \in \mathcal{C}(A, B)$
+    interchange equations for all \emph{0-cells} $A, B \in \mathcal{C}$ and all
+    \emph{2-cells} $\alpha, \beta, \gamma, \delta \in \mathcal{C}(A, B)$
   \begin{align}
-(\alpha \circ \beta) \bullet (\gamma \circ \delta) &= (\alpha \bullet \gamma)
-  \circ (\beta \bullet \delta) \\
+    (\alpha \circ \beta) \bullet (\gamma \circ \delta) &= (\alpha \bullet
+      \gamma) \circ (\beta \bullet \delta) \\
     (\alpha \otimes \beta) \bullet (\gamma \otimes \delta) &= (\alpha \bullet
@@
   \end{align}
+  Since $\otimes$ is a 2-functor, we have for all \emph{0-cells} $A, B \in
+    \mathcal{C}$ and all \emph{2-cells} $\alpha, \beta, \gamma, \delta \in
+    \mathcal{C}(A, B)$
+  \[
+    (\alpha \otimes \beta) \bullet (\gamma \otimes \delta) = (\alpha \bullet
+      \gamma) \bullet (\beta \bullet \delta)
+  \]
 \end{remark}
@@
   \begin{itemize}
-\item gates form the set of 1-cells 
-\item the 2-cells are generated by a finite set of \emph{rewriting} rules such
-  that
+    \item gates form the set of \emph{1-cells} 
+    \item the \emph{2-cells} are generated by a finite set of \emph{rewriting}
+      rules such that
     \begin{itemize}
       \item 2 rewriting rules with the same types must be equal
-\item each rewriting rule must be invertible, that is $\mathcal{C}(A, B)$ is a
-  \emph{groupoid} for all 0-cells $A, B$
+      \item each rewriting rule must be invertible, that is $\mathcal{C}(A, B)$
+        is a \emph{groupoid} for all \emph{0-cells} $A, B$
     \end{itemize}
@@
 
-2-cells are in general called \emph{rewrite sequence}, but if a rewrite sequence
-  has only 1 occurence of a rewrite rule, it is called an \emph{atomic rule}.
+  \emph{2-cells} are in general called \emph{rewrite sequence}, but if a rewrite
+    sequence has only 1 occurence of a rewrite rule, it is called an
+    \emph{atomic rule}.
 \end{definition}
@@
     \[
-\alpha = \alpha_1 \bullet \alpha_2 \bullet \alpha_3 \bullet \cdots \bullet \beta
+      \alpha = \alpha_1 \bullet \alpha_2 \bullet \alpha_3 \bullet \cdots \bullet
+        \beta
     \]
@@
     \begin{align}
-\alpha &= \alpha_1 \bullet \alpha_2 \bullet \alpha_3 \bullet \cdots \bullet
-  (\beta_1 \circ \beta_3) \bullet (\beta_2 \circ \beta_4) \cr
-\alpha &= \alpha_1 \bullet \alpha_2 \bullet \alpha_3 \bullet \cdots \bullet
-  \alpha_{i - 1} \bullet \alpha_i
+      \alpha &= \alpha_1 \bullet \alpha_2 \bullet \alpha_3 \bullet \cdots
+        \bullet (\beta_1 \circ \beta_3) \bullet (\beta_2 \circ \beta_4) \\
+      \alpha &= \alpha_1 \bullet \alpha_2 \bullet \alpha_3 \bullet \cdots
+        \bullet \alpha_{i - 1} \bullet \alpha_i
     \end{align}
-Otherwise we can always add an identity rewriting rule to get the interchange
-  law.
+    Otherwise we can always add an identity rewriting rule to get the
+      interchange law.
     \begin{align}
-\beta_1 \circ (\beta_3 \bullet \beta_4) &\Rightarrow (\beta_1 \bullet id) \circ
-  (\beta_2 \bullet \beta_3) \cr
-(\beta_1 \bullet \beta_2) \circ \beta_3 &\Rightarrow (\beta_1 \bullet \beta_2)
-  \circ (id \bullet \beta_3)
+      \beta_1 \circ (\beta_3 \bullet \beta_4) &\Rightarrow (\beta_1 \bullet id)
+        \circ (\beta_2 \bullet \beta_3) \\
+      (\beta_1 \bullet \beta_2) \circ \beta_3 &\Rightarrow (\beta_1 \bullet
+        \beta_2) \circ (id \bullet \beta_3)
     \end{align}
@@
   \begin{align}
-\alpha &= \alpha_1 \circ \alpha_2 \circ \alpha_3 \circ \cdots \circ \alpha_j \cr
+    \alpha &= \alpha_1 \circ \alpha_2 \circ \alpha_3 \circ \cdots \circ \alpha_j
+      \\
     \alpha &= \alpha_1 \otimes \alpha_2 \otimes \alpha_3 \otimes \cdots \otimes
@@
     \[
-(\alpha \otimes \beta) \circ (\gamma \otimes \delta) = (\alpha \circ \gamma)
-  \otimes (\beta \circ \delta)
+      (\alpha \otimes \beta) \circ (\gamma \otimes \delta) = (\alpha \circ
+        \gamma) \otimes (\beta \circ \delta)
     \]
@@
   \[
-\alpha_i = \alpha_{i, 1} \circ \alpha_{i, 2} \circ \alpha_{i, 3} \circ \cdots
-  \circ \alpha_{i, j}
+    \alpha_i = \alpha_{i, 1} \circ \alpha_{i, 2} \circ \alpha_{i, 3} \circ
+      \cdots \circ \alpha_{i, j}
   \]
@@
   \begin{align}
-x \oplus \bigvee y_i &= \bigvee x \oplus y_i \cr
+    x \oplus \bigvee y_i &= \bigvee x \oplus y_i \\
     \left(\bigvee x_i\right) \oplus y &= \bigvee x_i \oplus y
@@
   \begin{align}
-x \oplus x^{-1} &= e \cr
+    x \oplus x^{-1} &= e \\
     x^{-1} \oplus x &= e
@@
   \[
-\mathcal{E}(a, b) \oplus \mathcal{E}(a', b') \succeq \mathcal{E}(a \otimes a',
-  b \otimes b')
+    \mathcal{E}(a, b) \oplus \mathcal{E}(a', b') \succeq \mathcal{E}(a \otimes
+      a', b \otimes b')
   \]
@@
 \begin{definition}
-For all $(A, B) \in \mathcal{P}$ we define a homomorphism $V: \mathcal{C}(A, B)
-  \to \mathbb{Q}$, such that for all gates $f$
+  For all $(A, B) \in \mathcal{P}$ we define a homomorphism $V:
+    \mathcal{C}(A, B) \to \mathbb{Q}$, such that for all gates $f$
   \[
@@
   \begin{align}
-V(f \circ g) &\preceq V(f) \oplus V(g) \cr
+    V(f \circ g) &\preceq V(f) \oplus V(g) \\
     V(f \otimes g) &\preceq V(f) \oplus V(g)
@@
   \begin{align}
-V(\alpha \circ \beta) &= V(\alpha) \oplus V(\beta) \cr
-V(\alpha \otimes \beta) &= V(\alpha) \oplus V(\beta) \cr
+    V(\alpha \circ \beta) &= V(\alpha) \oplus V(\beta) \\
+    V(\alpha \otimes \beta) &= V(\alpha) \oplus V(\beta) \\
     V(\alpha \bullet \beta) &= V(\alpha) \oplus V(\beta)
@@
   \begin{align}
-V(\alpha: f \Rightarrow f) &= V(f) \oplus V(f)^{-1} \cr
+    V(\alpha: f \Rightarrow f) &= V(f) \oplus V(f)^{-1} \\
     &= e
@@
   that is
-\begin{align}
+  \[
     \forall f, V(id_f) = e
-\end{align}
+  \]
 \end{remark}
@@
 \begin{proposition}
-Given two rewriting sequence $\alpha, \beta: f \Rightarrow g$, then $V(\alpha) =
-  V(\beta)$.
+  Given two rewriting sequence $\alpha, \beta: f \Rightarrow g$, then $V(\alpha)
+    = V(\beta)$.
 
   \begin{proof}
-The set of all possible rewriting sequences is freely generated by the set of
-rewriting rules, $\bullet$, $\circ$ and $\otimes$ so both $\alpha$ and $\beta$
-  can be written
+    The set of all possible rewriting sequences is freely generated by the set
+      of rewriting rules, $\bullet$, $\circ$ and $\otimes$ so both $\alpha$ and
+      $\beta$ can be written
     \begin{align}
-\alpha &= \alpha_1 \odot \alpha_2 \odot \alpha_3 \odot \cdots \odot \alpha_n \cr
+      \alpha &= \alpha_1 \odot \alpha_2 \odot \alpha_3 \odot \cdots \odot
+        \alpha_n \\
       \beta &= \beta_1 \odot \beta_2 \odot \beta_3 \odot \cdots \odot \beta_m
@@
       $\beta_i$ are rewriting rule.
-If $n = 1$ then $\alpha$ is a rewriting rule, so $V(r) = V(f) \oplus V(g)^{-1}$.
+    If $n = 1$ then $\alpha$ is a rewriting rule, so $V(r) = V(f) \oplus
+      V(g)^{-1}$.
     Otherwise
     \begin{align}
-V(\alpha) &= V(\alpha_1 \odot \alpha_2 \odot \cdots \odot \alpha_n) \cr
-&= V(\alpha_1) \oplus V(\alpha_2) \oplus \cdots \oplus V(\alpha_n) \cr
+      V(\alpha) &= V(\alpha_1 \odot \alpha_2 \odot \cdots \odot \alpha_n) \\
+      &= V(\alpha_1) \oplus V(\alpha_2) \oplus \cdots \oplus V(\alpha_n) \\
       &= V(f \Rightarrow h_1) \oplus V(h_1 \Rightarrow h_2) \oplus \cdots \oplus
-  V(h_n \Rightarrow g) \cr
+        V(h_n \Rightarrow g) \\
       &= V(f) \oplus V(h_1)^{-1} \oplus V(h_1) \oplus V(h_2)^{-1} \oplus \cdots
@@
     \[
-V(f) \oplus V(h_1)^{-1} \oplus V(h_1) \oplus V(h_2)^{-1} \oplus V(h_2) \oplus
-  \cdots \oplus V(h_n) \oplus V(g)^{-1} = V(f) \oplus V(g)^{-1}
+      V(f) \oplus V(h_1)^{-1} \oplus V(h_1) \oplus V(h_2)^{-1} \oplus V(h_2)
+        \oplus \cdots \oplus V(h_n) \oplus V(g)^{-1} = V(f) \oplus V(g)^{-1}
     \]
@@
     \begin{cases}
-\textrm{if } f \Rightarrow g \in \mathcal{P}, &V(f \Rightarrow g) \cr
+      \textrm{if } f \Rightarrow g \in \mathcal{P}, &V(f \Rightarrow g) \\
       \textrm{otherwise } &\top
@@
     \begin{align}
-\mathcal{E}(f, g) &= V(f \Rightarrow g) \in \mathbb{Q} \cr
+      \mathcal{E}(f, g) &= V(f \Rightarrow g) \in \mathbb{Q} \\
       &= V(f) \oplus V(g)^{-1}
@@
     \begin{align}
-\mathcal{E}(f, f) &= V(id_f) \cr
-&= V(f \Rightarrow f)\cr
-&= V(f) \oplus V(f)^{-1} \cr
+      \mathcal{E}(f, f) &= V(id_f) \\
+      &= V(f \Rightarrow f)\\
+      &= V(f) \oplus V(f)^{-1} \\
       &= e
     \end{align}
-Finally, for all gates $(f, g, h)$ such that there exists $\alpha: f \Rightarrow
-  g$ and $\beta: g \Rightarrow h$ in $\mathcal{P}$, then the homomorphism $V$
-    give us
+    Finally, for all gates $(f, g, h)$ such that there exists $\alpha: f
+      \Rightarrow g$ and $\beta: g \Rightarrow h$ in $\mathcal{P}$, then the
+      homomorphism $V$ give us
     \begin{align}
-\mathcal{E}(f, h) &= V(f \Rightarrow h) \cr
-&= V(\alpha \circ \beta) \cr
-&= V(\alpha) \oplus V(\beta) \cr
-&= V(g \Rightarrow h) \oplus V(f \Rightarrow g) \cr
+      \mathcal{E}(f, h) &= V(f \Rightarrow h) \\
+      &= V(\alpha \circ \beta) \\
+      &= V(\alpha) \oplus V(\beta) \\
+      &= V(g \Rightarrow h) \oplus V(f \Rightarrow g) \\
       &= \mathcal{E}(g, h) \oplus \mathcal{E}(f, g)
@@
     become
-\[\begin{matrix}
-&\mathcal{E}(b, c) \oplus \top &\succeq \top \cr
-\textrm{or} &\top \oplus \mathcal{E}(a, b) &\succeq \top \cr
+    \[
+      \begin{matrix}
+        &\mathcal{E}(b, c) \oplus \top &\succeq \top \\
+        \textrm{or} &\top \oplus \mathcal{E}(a, b) &\succeq \top \\
         \textrm{or} &\top \oplus \top &\succeq \top
-\end{matrix}\]
+      \end{matrix}
+    \]
     which is still true.
@@
     \[
-V(\alpha: f \Rightarrow g) \oplus V(\beta: f' \Rightarrow g') = V(\alpha \otimes
-  \beta: f \otimes f' \Rightarrow g \otimes g')
+      V(\alpha: f \Rightarrow g) \oplus V(\beta: f' \Rightarrow g') = V(\alpha
+        \otimes \beta: f \otimes f' \Rightarrow g \otimes g')
     \]
@@
     \begin{align}
-\mathcal{E}(f \otimes g, g \otimes f) &= V(f \otimes g \Rightarrow g \otimes g)
-  \cr
-&= V(f \Rightarrow g \otimes g \Rightarrow f) \cr
-&\preceq V(f \Rightarrow g) \oplus V(g \Rightarrow f) \cr
-&= V(f) \oplus V(g)^{-1} \oplus V(g) \oplus V(f)^{-1} \cr
+      \mathcal{E}(f \otimes g, g \otimes f) &= V(f \otimes g \Rightarrow g
+        \otimes g) \\
+      &= V(f \Rightarrow g \otimes g \Rightarrow f) \\
+      &\preceq V(f \Rightarrow g) \oplus V(g \Rightarrow f) \\
+      &= V(f) \oplus V(g)^{-1} \oplus V(g) \oplus V(f)^{-1} \\
       &= e

### 2018-08-07 article.tex
...

--- 2018-08-06 article.tex
@@
 \begin{example}  \label{ex:quantum-circs}
-  TODO -- quanutm circuits
+  The main example of interest is \Circ, the PROP of quantum circuits.
+  Different authors, and different hardware manufacturers, prefer
+  different primitive gate sets but the following is typical:
+  \[
+  \begin{array}{cccccccccc}
+    \circZ{{}} & \qquad  & \circX{{}} & \qquad & \circH \\ 
+    \vstrut{1.2em}
+    Z : 1 \to 1 && X : 1 \to 1 && H : 1 \to 1\\
+    \\
+    \circbox{R_Z(\theta)} & \qquad & \circCPhase{\theta} & \qquad & \circCX  \\ 
+    \vstrut{1.2em}
+    R_Z(\theta) : 1 \to 1 &&  \CPhase\theta :2 \to 2 && \CX : 2 \to 2\\
+  \end{array}
+  \]
+  We omit the usual matrix interpretations but these can be found in
+  the standard textbook \cite{NieChu:QuantComp:2000}.  Note that in
+  this example we actually have an uncountable family of generating
+  gates, one for each value $\theta \in [0,2\pi)$.  Unlike the case of
+  Boolean circuits all quantum gates have the same number of inputs and
+  outputs.  In fact, \Circ is a \dag-PROP and all its but this will not concern
+  us here.
 \end{example}

### 2018-08-07 article.tex
minor edits

--- 2018-08-06 article.tex
@@
 
-\section{Mathematical Preliminaries}
+\section{Circuits and 2-PROPS}
 \label{sec:prelim}
@@
 \cite{Lack:2004sf}, however our interest here is to formalise a
-general notion of \emph{circuit} as a free PROP over a certain set of
-gates.
+general notion of \emph{circuit} made by wiring together certain
+atomic operations.  With this in mind, a PROP of circuits is
+henceforth a PROP freely generated from some monoidal signature; the
+members of the signature are called \emph{gates}, and a circuit is
+a morphism in such a PROP.
 
@@
   Boolean circuits all quantum gates have the same number of inputs and
-  outputs.  In fact, \Circ is a \dag-PROP, but this will not concern
+  outputs.  In fact, \Circ is a \dag-PROP and all its but this will not concern
   us here.
@@
 \begin{definition}\label{def:monoidal-two-cat}
-  A \emph{strict} monoidal 2-category $(\mathcal{C}, \otimes)$ is a strict
-    2-category with a 2-bifunctor $\otimes: \mathcal{C} \times \mathcal{C} \to
-    \mathcal{C}$.
-  \REM{This definition doesn't quite work without the additoin type
-    information on the \emph{2-cells}}
+  A \emph{strict} monoidal 2-category $(\mathcal{C}, \otimes)$ is a
+  strict 2-category with a 2-bifunctor $\otimes: \mathcal{C} \times
+  \mathcal{C} \to \mathcal{C}$.
+  \TODO{Denis: fix this definition}
+  \REM{What about the unit object?}
 \end{definition}
@@
 \begin{remark}[Interchange law]
+\TODO{DENIS : please fix this statement.  THe second and third clauses
+  are the same.!}
   Since $\circ$ and $\otimes$ are bifunctorial, we have the following
@@
 \end{remark}
+
+Putting the preceding definitions together, a 2-PROP is then a strict
+monoidal 2-category whose underlying monoidal 1-category is a PROP, or
+equivalently, a strict monoidal 2-category whose 0-cells are generated
+by a single object under $\otimes$.
 

### 2018-08-07 article.tex
minor edits
textrm to text
rewriting system example

@@
       $Id_f:f \Rightarrow f$.
-    \item For all \emph{0-cells} $a$, $b$, $c$ a bifunctor $\circ:
-      \mathcal{C}(b, c) \times \mathcal{C}(a, b) \to \mathcal{C}(a, c)$ called
+    \item For all \emph{0-cells} $A$, $B$, $C$ a bifunctor $\circ:
+      \mathcal{C}(B, C) \times \mathcal{C}(A, B) \to \mathcal{C}(A, C)$ called
       \emph{horizontal composition} of \emph{2-cells}.
@@
 \REM{Need here the name for a 2-cat with all \emph{2-cells} invertible;  the
-  term 2-groupoid refers to the case where both 1- and \emph{2-cells} are
+  term 2-groupoid refers to the case where both \emph{1-} and \emph{2-cells} are
   invertible.}
@@
 \begin{definition}\label{def:monoidal-two-cat}
-  A \emph{strict} monoidal 2-category $(\mathcal{C}, \otimes)$ is a
+  A \emph{strict} monoidal 2-category $(\mathcal{C}, \otimes, I)$ is a
   strict 2-category with a 2-bifunctor $\otimes: \mathcal{C} \times
-  \mathcal{C} \to \mathcal{C}$.
+  \mathcal{C} \to \mathcal{C}$ such that
+  \begin{itemize}
+    \item for all \emph{0-cells} $A$
+      \[
+        A \otimes I = A
+      \]
+    \item for all \emph{0-cells} $(A, B)$, $I$ is the monoidal unit of
+    $\mathcal{C}(A, B)$
+  \end{itemize}
   \TODO{Denis: fix this definition}
@@
   \[
-    (\alpha \otimes \beta) \bullet (\gamma \otimes \delta) = (\alpha \bullet
-      \gamma) \bullet (\beta \bullet \delta)
+    (\alpha \otimes \beta) \circ (\gamma \otimes \delta) = (\alpha \circ \gamma)
+      \otimes (\beta \circ \delta)
   \]
@@
 \label{sec:rewriting}
+
+As remarked earlier we don't include any equations in our notion of
+circuits
 
@@
 \begin{example}
-  TODO rewrite sequence \& atomic rule
+  Let's take the previous quantum circuit example \ref{ex:quantum-circs} and add
+    some \emph{noisy gates}, that is, gates which can failed with an $\epsilon$
+    probability.
+  \[
+    \begin{array}{cccccccccc}
+      \circZ{{\epsilon}} & \qquad  & \circX{{\epsilon}} & \qquad &
+        \circbox{H_\epsilon} \\ 
+      \vstrut{1.2em}
+      Z_\epsilon : 1 \to 1 && X_\epsilon : 1 \to 1 && H_\epsilon : 1 \to 1
+    \end{array}
+  \]
+  We form a 2-PROP by adding some rewriting rules (we omit the identies)
+  \[
+    \begin{array}{cccccccc}
+      \circZ{{\epsilon}} & \Rightarrow & \circZ{{}} & \alpha : Z_\epsilon \to Z \\
+      \vstrut{1.2em} \\
+      \circX{{\epsilon}} & \Rightarrow & \circX{{}} & \beta : X_\epsilon \to X \\
+      \vstrut{1.2em} \\
+      \circbox{H_\epsilon} & \Rightarrow & \circH & \gamma : H_\epsilon \to H
+    \end{array}
+  \]
+  Then, on the circuit
+  \[
+    \begin{array}{cccccccccc}
+      \circbox{H_\epsilon} \\ 
+      \vstrut{1.2em} \\
+      \circZ{{\epsilon}}
+    \end{array}
+  \]
+  a rewriting sequence can be $(id_{H_\epsilon} \otimes \alpha) \bullet (\beta
+    \otimes id_Z)$
+  \[
+    \begin{array}{cccccccccc}
+      \circbox{H_\epsilon} & \overset{id_{H_\epsilon}}{\Rightarrow} &
+        \circbox{H_\epsilon} & \overset{\beta}{\Rightarrow} & \circH \\ 
+      \vstrut{1.2em} \\
+      \circZ{{\epsilon}} & \overset{\alpha}{\Rightarrow} & \circZ{{}} &
+        \overset{id_Z}{\Rightarrow} & \circZ{{}} \\ 
+    \end{array}
+  \]
+  Moreover this sequence is atomic.
 \end{example}
@@
     \begin{cases}
-      \textrm{if } f \Rightarrow g \in \mathcal{P}, &V(f \Rightarrow g) \\
-      \textrm{otherwise } &\top
+      \text{if } f \Rightarrow g \in \mathcal{P}, &V(f \Rightarrow g) \\
+      \text{otherwise } &\top
     \end{cases}
@@
         &\mathcal{E}(b, c) \oplus \top &\succeq \top \\
-        \textrm{or} &\top \oplus \mathcal{E}(a, b) &\succeq \top \\
-        \textrm{or} &\top \oplus \top &\succeq \top
+        \text{or} &\top \oplus \mathcal{E}(a, b) &\succeq \top \\
+        \text{or} &\top \oplus \top &\succeq \top
       \end{matrix}

### 2018-08-08 article.tex
typo and remarks
better example
homomorphism examples

@@
       \item 2 rewriting rules with the same types must be equal
-      \item each rewriting rule must be invertible, that is $\mathcal{C}(A, B)$
+      \item each rewriting rule must be invertible, that is $\mathcal{P}(A, B)$
         is a \emph{groupoid} for all \emph{0-cells} $A, B$
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\newcommand{\circuitExample}[3]{\inline{%
+\begin{tikzpicture}[circuit]
+ \node (0) at (0, 0.5) {};
+ \node [box] (2) at (1, 0.5) {#1};
+ \node (4) at (3, 0.5) {};
+ \draw (0) to (2);
+ \draw (2) to (4);
+
+ \node (1) at (0, -0.5) {};
+ \node [box] (3) at (1, -0.5) {#2};
+ \node [box] (5) at (2, -0.5) {#3};
+ \node (6) at (3, -0.5) {};
+ \draw (1) to (3);
+ \draw (3) to (5);
+ \draw (5) to (6);
+\end{tikzpicture}
+}}
 \begin{example}
@@
     \begin{array}{cccccccc}
-      \circZ{{\epsilon}} & \Rightarrow & \circZ{{}} & \alpha : Z_\epsilon \to Z \\
+      \circZ{{\epsilon}} & \Rightarrow & \circZ{{}} & \alpha : Z_\epsilon
+        \Rightarrow Z \\
       \vstrut{1.2em} \\
-      \circX{{\epsilon}} & \Rightarrow & \circX{{}} & \beta : X_\epsilon \to X \\
+      \circX{{\epsilon}} & \Rightarrow & \circX{{}} & \beta : X_\epsilon
+        \Rightarrow X \\
       \vstrut{1.2em} \\
-      \circbox{H_\epsilon} & \Rightarrow & \circH & \gamma : H_\epsilon \to H
+      \circbox{H_\epsilon} & \Rightarrow & \circH & \gamma : H_\epsilon
+        \Rightarrow H
     \end{array}
@@
   \[
-    \begin{array}{cccccccccc}
-      \circbox{H_\epsilon} \\ 
-      \vstrut{1.2em} \\
-      \circZ{{\epsilon}}
-    \end{array}
+    \circuitExample{$H_\epsilon$}{$Z_\epsilon$}{$X$}
   \]
   a rewriting sequence can be $(id_{H_\epsilon} \otimes \alpha) \bullet (\beta
-    \otimes id_Z)$
+    \otimes id_Z) \circ (id \otimes id_X)$ \REM{I'm not sure of the sequence}
   \[
-    \begin{array}{cccccccccc}
-      \circbox{H_\epsilon} & \overset{id_{H_\epsilon}}{\Rightarrow} &
-        \circbox{H_\epsilon} & \overset{\beta}{\Rightarrow} & \circH \\ 
-      \vstrut{1.2em} \\
-      \circZ{{\epsilon}} & \overset{\alpha}{\Rightarrow} & \circZ{{}} &
-        \overset{id_Z}{\Rightarrow} & \circZ{{}} \\ 
-    \end{array}
+    \circuitExample{$H_\epsilon$}{$Z_\epsilon$}{$X$}
+    \xRightarrow{id_{H_\epsilon} \otimes \alpha}
+    \circuitExample{$H$}{$Z_\epsilon$}{$X$}
+    \xRightarrow{\beta \otimes id_Z}
+    \circuitExample{$H$}{$Z$}{$X$}
   \]
@@
   For all $(A, B) \in \mathcal{P}$ we define a homomorphism $V:
-    \mathcal{C}(A, B) \to \mathbb{Q}$, such that for all gates $f$
+    \mathcal{P}(A, B) \to \mathbb{Q}$, such that for all gates $f$
   \[
     V(f) \in \mathbb{Q}
-  \]
-  For all rewriting rules $\alpha: f \Rightarrow g$
-  \[
-    V(\alpha) = V(f) \oplus V(g)^{-1}
   \]
@@
   \end{align}
+
+  We all required that for all rewriting rules $\alpha: f \Rightarrow g$
+  \[
+    V(\alpha) = V(f) \oplus V(g)^{-1}
+  \]
+
 \end{definition}
@@
     \oplus V(g)^{-1}$.
-  Note that the definition of $V$ gives that only for rewriting rules.
+  Note that the homomorphism $V$ is now well defined
 \end{corollary}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%        HOMOMORPHISM EXAMPLES        %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{example}[Gates counter]
+  Let's start with a simple example, a gates counter homomorphism $N$ over the
+    quantale the additive group of natural numbers.
+  \begin{align*}
+    N(f) & = 1 \\
+    N(f \circ g) &= N(f) + N(g) \\
+    N(f \otimes g) &= N(f) + N(g)
+  \end{align*}
+  On rewriting sequences we have
+  \[
+    N(f \Rightarrow g) = N(f) - N(g)
+  \]
+  Then the induced $\mathbb{Q}$-enriched category become a tool which give us
+    the number of gates added or removed given a rewriting sequence.
+\end{example}
+
+\begin{example}[Depth \& time counters]
+  Two similar counters can be made using the depth and the time of a circuit
+  \begin{align*}
+    D(f) &= 1 \\
+    D(f \circ g) &= D(f) + D(g) \\
+    D(f \otimes g) &= \max(D(f), D(g))
+  \end{align*}
+  and
+  \begin{align*}
+    T(f) &= t_f \\
+    T(f \circ g) &= T(f) + T(g) \\
+    T(f \otimes g) &= \max(T(f), T(g))
+  \end{align*}
+  With $t_f$ defined for the generators of the PROP.
+  It is easily to check that
+  \begin{align*}
+    \max(D(f), D(g)) \preceq D(f) + D(g)
+    \max(T(f), T(g)) \preceq T(f) + T(g)
+  \end{align*}
+  as required by the homomorphism.
+\end{example}
+
 \section{Conclusions and further work}

### 2018-08-08 article.tex
typo

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{example}[Gates counter]
+\begin{example}[Gates count]
   Let's start with a simple example, a gates counter homomorphism $N$ over the
@@
 
-\begin{example}[Depth \& time counters]
+\begin{example}[Depth \& time count]
   Two similar counters can be made using the depth and the time of a circuit
@@
   \begin{align*}
-    \max(D(f), D(g)) \preceq D(f) + D(g)
+    \max(D(f), D(g)) \preceq D(f) + D(g) \\
     \max(T(f), T(g)) \preceq T(f) + T(g)
